### 2023-05-05 freechains/freechains.tex
...

\documentclass[12pt]{article}

\usepackage{sbc-template}
\usepackage{graphicx,url}
\usepackage[utf8]{inputenc}

\usepackage{verbatim}
\usepackage{xspace}
\newcommand{\FC}       {Freechains\xspace}
\newcommand{\reps}     {\emph{reps}\xspace}
\newcommand{\onerep}   {\emph{1~rep}\xspace}
\newcommand{\nreps}[1] {\emph{#1~reps\xspace}}
\newcommand{\code}[1]  {\texttt{\footnotesize{#1}}}
\newcommand{\Xon} {$1{\rightarrow}N$\xspace}
\newcommand{\Xno} {$1{\leftarrow}N$\xspace}
\newcommand{\Xnn} {$N{\leftrightarrow}N$\xspace}
\newcommand{\Xoo} {$1{\leftrightarrow}1$\xspace}
\newcommand{\Xo}  {$1{\hookleftarrow}$\xspace}

\renewcommand{\theenumi}{\alph{enumi}}

\hyphenation{off-line}

\sloppy

\title{
    Peer-to-Peer Permissionless Consensus via Authoring Reputation
}

\author{Anonymous}
\address{Anonymous}

\begin{document}

\maketitle

\begin{abstract}
Public Internet forums suffer from excess and abuse, such as SPAM and fake
news.
Centralized platforms employ filtering and anti-abuse policies, but imply full
trust from users.
%
We propose a permissionless Sybil-resistant peer-to-peer protocol for content
sharing.
Our main contribution is a reputation system that moderates content and, at the
same time, delivers network consensus.
We can trace a parallel with Bitcoin as follows:
    consolidated posts create reputation (vs proof-of-work),
    likes and dislikes transfer reputation (vs transactions), and
    aggregate reputation determines consensus (vs longest chain).
%
The reputation mechanism depends exclusively on the human authoring ability
(\emph{proof-of-authoring}), which is slow and scarce, thus suitable to
establish consensus.
\end{abstract}

\begin{resumo} 
Fórums públicos de Internet sofrem de excesso e abuso, tais como SPAM e
notícias falsas.
Plataformas centralizadas adotam políticas de filtragem e anti abuso, mas
exigem confiança dos seus usuários.
%
Nós propomos um protocolo peer-to-peer não permissionado para compartilhamento
de conteúdo que é resistente a ataques Sybil.
Nossa principal contribuição é um sistema de reputação que serve para moderar o
conteúdo e ao mesmo tempo garantir consenso na rede.
%
Nós traçamos um paralelo com o Bitcoin da seguinte forma:
    postagens consolidadas criam reputação (vs \emph{proof-of-work}),
    \emph{likes} e \emph{dislikes} transferem reputação (vs transações), e
    a reputação agregada determina o consenso (vs cadeia mais longa).
%
O mecanismo de reputação depende exclusivamente da habilidade de autoria humana
(\emph{proof-of-authoring}), que é lenta e escassa, e portanto adqueada para
estabelecer consenso.
\end{resumo}

\section{Introduction}
\label{sec.introduction}

Content publishing in Internet forums and social media is
increasingly more centralized in a few companies (e.g. Facebook and
Twitter)~\cite{internet.fixing,p2p.osn,p2p.dosn}.
%These companies benefit from closed protocols and network effects to keep
%users locked in their platforms.
On the one hand, these companies offer free storage, friendly user interfaces,
and robust access.
On the other hand, they concentrate power by collecting users' data and
``algorithmizing'' consumption.
%
Peer-to-peer alternatives~\cite{p2p.survey} eliminate intermediaries, but
strive to achieve consistency while dealing with malicious users.
%given the decentralization of authority and infrastructure.
% and push to end users the responsibility to manage data and connectivity.

In an ideal Internet forum, all messages or posts
(i)   reach all users,
(ii)  are delivered in a consistent order, and
(iii) are respectful and on topic.
In a centralized system, items (i) and (ii) are trivially achieved assuming
availability and delivery order in the service, while for item (iii), users
have to trust the service to moderate content.
In a decentralized setting, however, none of these demands are easily
accomplished.
A common approach in gossiping protocols is to proactively replicate and
disseminate posts among peers until they reach all
users~\cite{p2p.survey,p2p.byz}.
However, this approach does not guarantee consensus since posts can be received
in conflicting orders~\cite{p2p.intention,p2p.dvcs}.
%As an example, antagonistic messages such as \emph{"X is final"} vs
%\emph{"Y is final"} might be sent concurrently, preventing the network to
%determine as a group its intention as \emph{X} or \emph{Y}.
%
Consensus is key to eradicate Sybil attacks~\cite{p2p.sybil}, which are the
major threat to decentralized applications in general: without consensus, it
is not possible, \emph{at the protocol level}, to distinguish between correct
and malicious users in order to satisfy item (iii).
%fundamental in collaborative applications, such as open discussion
%forums, chat rooms, and trading services (e.g., delivery \& auction).
%Otherwise, they become impractical due to excess and abuse (e.g., SPAM and
%illegal content).
%At the very core, consensus is

\begin{comment}
Consensus is notably challenging to the point that decentralized protocols
partially abdicate of it.
%For instance, it is typically either possible to have
%\emph{multi-user/single-node} or \emph{single-user/multi-node}, but not
%\emph{multi-user/multi-node} consensus:
%
On the one hand, federated protocols~\cite{p2p.ecosystem} offer
\emph{multi-user/single-node consensus}, in which multiple users can exchange
messages consistenlty within a single trusted server, but not globally across
multiple servers.
%
On the other hand, a protocol like Scuttlebutt~\cite{p2p.scuttlebutt} offers
\emph{single-user/multi-node consensus}, in which a single user has full
authority over its own content across machines, but multiple users cannot reach
consensus even in a local machine.
%
Our goal is to provide \emph{multi-user/multi-node consensus} (just
\emph{consensus}, from now on) in the context of decentralized content sharing.
\end{comment}

Bitcoin~\cite{p2p.bitcoin} is the first permissionless protocol to resist
Sybils through consensus.
Its key insight is to rely on a scarce resource---the \emph{proof-of-work}---to
establish consensus.
%The protocol maintains a single dispendious timeline consisting of linked
%blocks with transactions, which represent the consensus.
%Alternative timelines need to spend more resources to substitute the consensus.
The protocol is Sybil resistant because it is expensive to write to its
unique timeline (either via proof-of-work or transaction fees).
%
However, Bitcoin and cryptocurrencies in general are not suitable for content
sharing because
    (i)   they enforce a unique timeline to preserve value and immunity to
          attacks;
    (ii)  they lean towards concentration of power due to scaling effects; and
    (iii) they impose an external economic cost to use the protocol.
These issues threaten our original decentralization goals.
In particular, a unique timeline implies that all Internet content should be
subject to the same consensus rules, which neglects all subjectivity that is
inherent to social content.
%
Another limitation of cryptocurrencies is that it is not possible to revoke
content in the middle of a blockchain, which is inadmissible considering
illegal content (e.g., hate speech).

In this work, we adapt the idea of scarce resources to reach consensus, but in
the context of social content sharing.
%
Our first contribution is to recognize the actual published contents as the
protocol scarce resources, since they require human work.
%
Work is manifested as new posts, which if approved by others, reward authors
with reputation tokens, which are used to evaluate other posts with likes and
dislikes.
With such \emph{proof-of-authoring} mechanism, token generation is expensive,
while verification is cheap and made by multiple users.
%
Due to decentralization, posts in a timeline form a causal graph with only
partial order, which we promote to a total order based on the reputation of
authors.
The consensus order is fundamental to detect conflicting operations, such as
likes with insufficient reputation (akin to Bitcoin's double spending).
%
Our second contribution is to allow that users create diversified forums of
interest (instead of a singleton blockchain), each counting as an independent
timeline with its own subjective consensus etiquette.
%
Our third contribution is to support content removal without compromising the
integrity of the decentralized blockchain.
Users have the power to revoke posts with dislikes, and peers are forced to
remove payloads, only forwarding associated metadata.
%
We integrated the proposed consensus algorithm into \FC~\cite{fcs.sbseg20}, a
practical peer-to-peer content dissemination protocol that provides strong
eventual consistency~\cite{p2p.crdts,p2p.sec}.
%
To show the practicability of the consensus mechanism, we simulated months of
activity of a chat channel and years of a newsgroup forum, both extracted from
publicly available Internet archives.

The rest of the paper is organized as follows:
In Section~\ref{sec.design}, we describe the design of the reputation and
consensus mechanism for public forums.
In Section~\ref{sec.freechains}, we describe the concrete reputation rules we
implemented for \FC.
In Section~\ref{sec.evaluation}, we evaluate the performance of the protocol
in real-world public forums.
In Section~\ref{sec.related}, we compare our system with publish-subscribe
protocols, federated applications, and fully peer-to-peer systems.
In Section~\ref{sec.conclusion}, we conclude this work.

\newpage
\section{The Reputation and Consensus Mechanism}
\label{sec.design}

In the absence of moderation, permissionless peer-to-peer forums are
impractical, mostly because of Sybils abusing the system.
For instance, it should take a few seconds to generate thousands of fake
identities and SPAM millions of messages into the system.
%Hence, without moderation, there are no limits on the number and size of posts
%and no reasonable policy to distinguish quality.
For this reason, we propose a reputation system that works together with a
consensus algorithm to resist Sybil attacks.
% and make peer-to-peer public forums practical.

We propose a reputation system in which users can spend tokens named \reps to
post and rate content in the forums:
a \code{post} initially penalizes authors until it consolidates and counts
positively;
a \code{like} is a positive feedback that helps subscribers to distinguish
content amid excess;
a \code{dislike} is a negative feedback that revokes content when crossing a
threshold.
Table~\ref{fig.general} summarizes the reputation operations and their goals.
To prevent Sybils, users with no \reps cannot perform these operations,
requiring a welcoming like from any other user already in the system.
The fact that likes are zero-sum operations, which only transfer reputation,
eliminates the incentives from malicious users to invite Sybils into the
system.
The only way to generate new \reps is to post content that other users approve,
which demands non-trivial work resistant to automation.%
\footnote{
    Note that to some extent work from AI may be indistinguishable from humans,
    but its frequency and relevance are still subject to human evaluation,
    which is subjective and evolves over time.
}

\begin{table}
\centering
\includegraphics[width=0.75\textwidth]{general.png}
\caption{General reputation operations in public forums.}
\label{fig.general}
\end{table}

Bitcoin employs proof-of-work to mitigate Sybil attacks.
However, CPU or other extrinsic resources are not evenly distributed among
humans, specially in communications using battery-powered devices.
%
Instead, considering the context of public forums, we can take advantage of the
human authoring ability as an intrinsic resource.
Creating new content is hard and takes time, but is comparatively easy to
verify and rate.
Therefore, in order to impose scarcity, we determine that only content
authoring generates \reps, while likes and dislikes just transfer \reps between
users.
%but limited by periods of time to prevent excess (e.g., once in a day per user).
%Additionally,
%
Nevertheless, scarce operations are not yet sufficient because they demand
consensus to establish an order in time across the network to prevent
inconsistent operations.
As an example, consider a malicious author with a single unit of \reps posting
new messages using multiple peers at the same time.
According to the \emph{Expense} rule of Table~\ref{fig.general}, only one of
these messages should be accepted.
However, without consensus, it is not possible to globally determine which
message to accept, since each peer would supposedly accept the first message it
sees.
On the one hand, in order to validate operations consistently, we need the same
message ordering across all peers in the network.
On the other hand, we typically can only represent peer-to-peer public forums
as DAGs with causal relationships between messages, which provides at most
partial order.

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{reps2.png}
\caption{
    (A) A public forum DAG with a common prefix and two branches.
    (B) Total order between blocks of the DAG after consensus.
}
\label{fig.reps}
\end{figure}

Therefore, our key idea to stablish consensus in forum DAGs is to favor forks
with posts from users that constitute the majority of the reputation in the
network.
These forks have more associated work from active users and are analogous to
longest chains in Bitcoin.
%
In technical terms, we can adapt a topological sorting algorithm to favor
reputation when deciding between branches in a forum DAG.

Figure~\ref{fig.reps}.A illustrates the consensus criteria.
%, still abstractly, since we did not discuss the actual rules for content creation and rating.
A public forum DAG has a common prefix with signed posts from users $a$, $b$,
and $c$.
Let's assume that within the prefix, users $a$ and $b$ have contributed with
better content and have more reputation combined than $c$ has alone (i.e.,
$8+5>3$).
%
After the prefix, the forum forks in two branches:
in \code{branch-1}, only user $c$ remains active and we see that new users $x$
and $y$ (with no previous reputation in the common prefix) generate a lot of
new content;
in \code{branch-2}, only users $a$ and $b$ participate but with less activity.
Nevertheless, \code{branch-2} would be ordered first in this case because,
before the forking point, $a$ and $b$ have more reputation than $c$, $x$, and
$y$ combined.
%
User $c$ here might represent a malicious user trying to cultivate fake
identities $x$ and $y$ in separate of the network during weeks to accumulate
\reps.
%However, the whole malicious \code{branch-1} is vulnerable because users in
%\code{branch-2} with more previous reputation take the priority and can
%overthrow user $c$.

Figure~\ref{fig.reps}.B indicates the resulting consensus order between blocks
in the forum.
All operations in \code{branch-2} appear before any operation in
\code{branch-1}.
%
At any point in the consensus timeline, if an operation fails, all remaining
blocks in the offending branch are removed from the primary DAG.
As an example, suppose that the last post by $a$ (in gray) is a dislike to user
$c$.
Then, it's possible that the last post by $c$ (in red), now (suppose) with
\nreps{0}, is rejected together with all posts by $y$ and $x$ in sequence.
%
The removal rule for the remaining blocks is a direct consequence of tamper
proof Merkle~DAGs~\cite{p2p.ipfs,p2p.bitcoin} used in permissionless protocols,
which cannot support graph modifications.

Note that the consensus order exists only for accountability purposes, and is
a view of the primary DAG structure.
%
Note also that users in the branch with more reputation can react to attacks
even after the fact.
For instance, users $a$ and $b$ can pretend that they did not yet see
\code{branch-1} and post extra dislikes to user $c$ from \code{branch-2} so
that a further merge removes all blocks of \code{branch-1} from the DAG.
%
There are some other relevant considerations about forks and merges:
%
Peers that first received branches with less reputation will need to reorder
all blocks starting at the forking point.
This might involve removing content in the end-user software.
This behavior is similar to Bitcoin's blockchain reorganization, when a peer
detects a new longest chain. % and disconsiders old blocks.
%
Likewise, peers that first saw branches with more reputation just need to put
the other branch in sequence with no reordering at all.
%and do not need to recompute anything.
This behavior is expected to happen in the majority of the network.
%
Note that unlike Bitcoin, forks are not only permitted but encouraged due to
the local-first software principle~\cite{p2p.local}, in which networked
applications can work locally while offline.
However, the longer a peer remains disconnected, the more conflicting
operations it may see, and the higher are the chances of rejection when
rejoining.

As a counterpoint to the consensus order in Figure~\ref{fig.reps}.B, maybe
users $a$ and $b$ have abandoned the forum for months, and thus \code{branch-1}
is actually legit.
In this case, users $a$ and $b$ might be the ones trying to take over the
forum.
Yet another possibility is that both branches are legit but became disconnected
for a long period of time.
In any case, it is unacceptable that a very old remote branch affects a long
active local forum.
%
For this reason, the consensus algorithm includes an extra constraint that
prevents long-lasting local branches to merge, creating \emph{hard forks} in
the network.
A hard fork occurs when a local branch crosses a predetermined and irreversible
threshold of \emph{7 days} or \emph{100 posts} of activity.
In this case, regardless of the remote branch reputation, the local branch
takes priority and is ordered first.
This situation is analogous to a hard fork in Bitcoin and the branches will
never synchronize again.
More than simply numeric disputes, hard forks represent social conflicts in
which reconciling branches is no longer possible.
%
Figure~\ref{fig.hard} illustrates hard forks by distinguishing \emph{stable
consensus}, which cannot be reordered, from \emph{unstable consensus}, which may
still be affected by incoming branches.
The activity threshold counts backwards, from the latest local block in the
unstable consensus.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{n100-d7.png}
\caption{
    Stable consensus freezes the order of blocks once they cross the threshold.
    The unstable order may still be affected by incoming branches.
}
\label{fig.hard}
\end{figure}

In summary, the rules to merge a branch from a remote machine \code{j} into a
branch from a local machine \code{i} are as follows:
\begin{itemize}
    \item \code{i} is first if it crosses the activity threshold of
          \emph{7 days} or \emph{100 posts}, regardless of \code{j}.
    \item \code{i} or \code{j} is first, whichever has more reputation in the
          common prefix.
    \item otherwise, branches are ordered by an arbitrary criteria, such as
          lexicographical order of the block hashes immediately after the
          common prefix.
\end{itemize}

\begin{figure}
\centering
\includegraphics[width=0.75\textwidth]{state-revoked.png}
\caption{
    State machine of posts:
    \code{BLOCKED} posts are not linked in the DAG.
    \code{ACCEPTED} posts are linked and retransmitted.
    The payload of \code{REVOKED} posts are not retransmitted.
}
\label{fig.state}
\end{figure}

Finally, we consider that revoking posts is fundamental in the context of
social content publishing, and thus, an important contribution of this work.
As described in Figure~\ref{fig.state}, a forum post has three possible states:
\code{BLOCKED}, \code{ACCEPTED}, or \code{REVOKED}.
%
If the author has reputation, a new post is immediately \code{ACCEPTED} in the
forum.
%
Otherwise, it is \code{BLOCKED} and requires a like from another user.
Blocked posts are not considered part of the forum DAG in the sense that new
posts do not link back to it.
In addition, peers are not required to hold blocked posts and neither
retransmit them to other peers.
However, if blocked posts are not disseminated, new users will never have the
chance to be welcomed with a like.
A reasonable policy is to hold blocked posts in a temporary bag and retransmit
them for some visibility in the network.
%
Once accepted, a post becomes part of the forum and can never be removed
again, since Merkle~DAGs are immutable by design.
%Note that blocked posts that become accepted are always succeeded by a
%\emph{like} (Figure~\ref{fig.forum}).
%
However, if the number of dislikes exceeds an arbitrary threshold (e.g., more
dislike than likes), the block becomes \code{REVOKED} and its payload is not
retransmitted to other peers.
Note that a block hash does not depend on its associated payload, but only on
the payload hash.
Hence, it is safe to remove the payload as long as one can prove its revoked
state.
Later, if the post receives new likes, it means that the payload is still known
somewhere and peers can request it when synchronizing again.

\section{Public Forums in Freechains}
\label{sec.freechains}

\FC~\cite{fcs.sbseg20} is an unstructured peer-to-peer topic-based
publish-subscribe protocol, in which each \emph{chain} is a replicated
Merkle-DAG representing the causal relationships between the messages.
%
The protocol operation is typical of publish-subscribe systems: an author
publishes a post to a chain, and subscribed users eventually receive the
message.
% by synchronizing DAGs.

\begin{table}
\centering
\includegraphics[width=\textwidth]{rules.png}
\caption{
    Reputation rules for public forum chains in \FC.
    The chosen constants ($30~reps$, $24h$, etc) are arbitrary and target
    typical Internet forums.
}
\label{fig.rules}
\end{table}

We integrated the proposed reputation and consensus mechanism with \FC to
support content moderation and mitigate abuse.
Table~\ref{fig.rules} details the rules we conceived, which are discussed as
follows through a concrete example.
Authors have to sign posts in order to be accounted by the reputation system
and operate in the chains.
We start by creating an identity whose public key is assigned as the pioneer in
a chain named \emph{forum}:

{\footnotesize
\begin{verbatim}
 > freechains keys pubpvt 'pioneer-password'
 4B56AD.. DA3B5F..  <-- public and private keys
 > freechains 'forum' join '4B56AD..'
 10AE3E..           <-- hash representing the chain
 > freechains 'forum' post --sign='DA3B5F..' \
    'The purpose of this chain is...'
 1_CC2184..         <-- hash representing the post
\end{verbatim}
}

The \code{join} command in rule~\code{1.a} bootstraps a public chain,
assigning \nreps{30} equally distributed between an arbitrary number of
pioneers indicated through their public keys.
The pioneers shape the initial culture of the chain with the first posts and
likes, while they gradually transfer \reps to other authors, which also
transfer to other authors, expanding the community.
%
The \code{post} command in sequence is signed by the single pioneer (in this
example) and indicates the purpose of the chain for future users.

The most basic concern in public forums is to resist Sybils abusing the
chains.
Fully peer-to-peer systems cannot rely on logins or {\footnotesize CAPTCHAs}
due to the lack of a central authority.
Viable alternatives include (i) building social trust graphs, in which users
already in the community vouch for new users, or (ii) imposing explicit costs
for new posts, such as proof of work.
%
We propose a mix between trust graphs and economic costs.
%
Rule~\code{4.a} imposes that authors require at least \onerep to post,
effectively blocking Sybil actions.
To vouch for new users, rule~\code{3.a} allows an existing user to like a
newbie's post to unblock it, but at the cost of \onerep.
This cost prevents that malicious members unblock new users indiscriminately,
which would be a breach for Sybils.
For the same reason, rule~\code{2} imposes a temporary cost of \onerep for
each new post.
%
Note that the pioneers rule~\code{1.a} solves the chicken-and-egg problem
imposed by rule~\code{4.a}: if new authors start with no \reps, but require
\reps to operate, it is necessary that some authors have initial \reps to boot
the chains.

In the next sequence of commands, a new remote user joins the same public chain
and posts a message, which is welcomed with a like signed by the pioneer:

{\footnotesize
\begin{verbatim}
 > freechains keys pubpvt 'newbie-password'
 503AB5.. 41DDF1..  <-- public and private keys
 > freechains 'forum' join '4B56AD..'
 10AE3E..           <-- same pioneer as before
 > freechains 'forum' post 'Im a newbie...' \
    --sign='41DDF1..'
 2_C3A40F..         <-- blocked post
 > freechains 'forum' like '2_C3A40F..' \
    --sign='DA3B5F..'
 3_59F3E1..         <-- hash representing the like
\end{verbatim}
}

Note that chains with the same name but different pioneers would be
incompatible because the hash of genesis blocks depend on the pioneers' public
keys.

Figure~\ref{fig.forum} illustrates the chain DAG up to the like operation.
The pioneer starts with \nreps{30} (rule~\code{1.a}) and posts the initial
message.
%
New posts penalize authors with \nreps{-1} during at most 12 hours
(rule~\code{2}), which depends on the activity succeeding (and including) the
new post.
The more activity from reputed authors, the less time the discount persists.
In the example, since the post is from the pioneer controlling all \reps in the
chain, the penalty falls immediately and she remains with \nreps{30}.
This mechanism limits the excess of posts in chains dynamically.
For instance, in slow technical mailing lists, it is more expensive to post
messages in sequence.
However, in chats with a lot of active users, the penalty can decrease to zero
quickly.

\begin{figure}
\centering
\includegraphics[width=0.75\textwidth]{forum.png}
\caption{
    The \code{like} approves the newbie message into the \code{\#forum} DAG.
}
\label{fig.forum}
\end{figure}

Back to Figure~\ref{fig.forum}, a new user with \nreps{0} tries to post a
message (hash~\code{C3A40F..}) and is blocked (rule~\code{4.a}), as the red
background highlights.
But the pioneer liked the blocked message, decreasing herself to \nreps{29}
and increasing new user to \onerep (rule~\code{3.a}).
Note that the newbie post is not penalized (rule~\code{2}) because it is
followed by the pioneer like, which still controls all \reps in the chain.

With no additional rules to generate \reps, the initial \nreps{30} would
constitute the whole ``chain economy'' forever.
For this reason, rule~\code{1.b} rewards authors of new posts with \onerep,
but only after 24 hours.
This rule stimulates content creation and grows the economy of chains.
The 24-hour period gives sufficient time for other users to judge the post
before rewarding the author.
It also regulates the growth speed of the chain.
In Figure~\ref{fig.forum}, after 1 day, the pioneer would now accumulate
\nreps{30} and the new user \nreps{2}, growing the economy in \nreps{2} as
result of the two consolidated posts.
Note that rule~\code{1.b} rewards at most one post of each author at a time.
Hence, new posts during the 24-hour period will not reward each of them with
extra \reps.
Note also that rule~\code{4.b} limits authors to at most \nreps{30}, which
provides incentives to spend likes and thus decentralize the network, while
rule~\code{4.c} limits the size of posts to at most \emph{128kB} to prevent
DDoS attacks using gigantic blocked posts.

Likes and dislikes (rules \code{3.a} and \code{3.b}) serve three purposes
in the chains:
    (i) welcoming new users,
    (ii) measuring the quality of posts, and
    (iii) revoking abusive posts (SPAM, fake news, illegal content, etc).
%
The quality of posts is subjective and is up to users to judge them with likes,
dislikes, or simply abstaining.
%
This way, access to chains is permissionless in the sense that the actual peers
and identities behind posts are not directly assessed by the protocol, but
instead by the other users in the system.
%
The reputation of a given post is the difference between its likes and
dislikes, which can be used in end-user software for filtering and highlighting
purposes.
%
On the one hand, since \reps are finite, users need to ponder to avoid
indiscriminate expenditure.
On the other hand, since \reps are limited to at most \nreps{30} per author
(rule~\code{4.b}), users also have incentives to rate content.
Hence, these upper and lower limits work together towards the quality of the
chains.
%
Note that a dislike shrinks the chain economy since it removes \reps from both
the origin and target.
Finally, the actual contents of a post may be revoked if it has at least 3
dislikes, and more dislikes than likes (rule~\code{3}).
However, considering that \reps are scarce, dislikes are encouraged to combat
abusive behavior, but not to eliminate divergences of opinion.

\section{Experiments with Public Forums}
\label{sec.evaluation}

In this section, we perform experiments to evaluate the performance and
practicability of the protocol.
%
As detailed further, we measure the following evaluation parameters:
    (a) metadata overhead,
    (b) consensus runtime,
    (c) graph forks, and
    (d) blocked messages.

We simulate the behavior of two publicly available forums as if they were using
\FC:
%
    a chat channel from the Wikimedia Foundation%
\footnote{ Chat: \url{https://archive.org/download/WikimediaIrcLogs/} }, and
    the \emph{comp.compilers} newsgroup%
\footnote{ Newsgroup: \url{https://archive.org/download/usenet-comp} }.
%
Chats and newsgroups represent typical public forums with
    faster interactions with shorter payloads (chats), and
    slower interactions with larger payloads (newsgroups).
%
We only simulate the first \emph{10.000} messages of the forums, which
represent 3 months of activity in the chat and 9 years in the newsgroup.

The simulation spawns $N$ peers, each joining the same chain with the same
arguments.
For each message in the original forum, we
    (i)   set the timestamp of all peers to match the date,
    (ii)  create a pair of keys if the author is new,
    (iii) post the message from a random peer in $N$,
    (iv)  like the post if the author has no reputation, and
    (v)   synchronize the chain with $M$ random peers.
%
Since all messages are part of the original archive, we always perform step
(iv) to unblock messages from newbies, which we use to measure the evaluation
parameter (d).
Step (v) will inevitably create forks in the chain for any \texttt{M<N}, which
we measure in evaluation parameter (c).

For the newsgroup, we use \texttt{N=15} and \texttt{M=5}, which represents a
larger number of peers with few interconnections to stress the local-first
nature of the protocol.
For the chat, we use \texttt{N=5} and \texttt{M=3}, which represents a smaller
number of peers with more interconnections.
%
We executed each simulation $4$ times in a conventional desktop PC (\texttt{i7}
CPU, \texttt{8GB} RAM, \texttt{512GB} SSD).
Since the variations were negligible, we always discuss the median measures.

\begin{comment}
We assume chats are less distributed given their real-time requirements, i.e.,
users would connect to a few but more interconnected peers.
Note that \FC focus on the decentralization of authority, which is not directly
related to network distribution, which may vary from applications.
\end{comment}

Evaluation item (a) measures the protocol overhead due to blockchain metadata,
which consists of a timestamp, author signature, and hashes (block id, payload,
backlinks, and likes).
%
The original chat archive is \texttt{800kB} in size, or \texttt{80B} for each
message, which includes a timestamp, a username, and the actual payload.
The simulated chat chain is \texttt{8MB} in size, which indicates a $10x$
overhead.
% (\code{20100825 [03:37:43]})
% (\code{hello everyone})
% (\code{<Ashlee>})
The original newsgroup is \texttt{30MB} in size, or \texttt{3kB} for each
message, which includes a timestamp, a sender, a subject, and the actual
payload (typically much longer).
The simulated newsgroup chain is \texttt{42MB} in size, which indicates a
$50\%$ overhead.
% (\code{29 Aug 88 23:36:58 GMT})
% (\code{mark@ramid.com})
% (\code{Error messages and Yacc})
%
It is clear that the metatada overhead is not negligible, specially for short
chat messages, but decreases as the payload increases.

Evaluation item (b) accounts the consensus algorithm applied locally at each
chain.
We measured the time to sort blocks in a local DAG both for the first time
(without any caches), and incrementally (from stable consensus caches).
%
For the chat chain, it takes \texttt{125s} and \texttt{50ms} for the initial
and incremental sorts respectively, while for the newsgroup chain, it takes
\texttt{100s} and \texttt{70ms}.
%
%Currently, we synchronize the DAGs takes hours because we the synchronization time
%
The incremental sort is limited to 7 days or 100 posts, regardless of the size
of the chain, which conveniently settles an upper bound on the input size of
the consensus algorithm.
%
Given that we use plain JSON files in the file system, we consider an
incremental consensus under \texttt{100ms} to be a practical upper bound.

Evaluation item (c) counts the number of forks in chain DAGs, which indicates
the level of asynchrony between peers following the local-first principle
(supposedly).
We calculate the ratio of forks over the total number of messages, e.g., a DAG
with $100$ messages and $10$ forks has a ratio of $10\%$.
%Based on the choices for $M$ and $N$ for each forum, we expect the newsgroup to
%have a higher ratio, since peers synchronize less.
We found a ratio of $18\%$ for the chat and $14\%$ for the newsgroup, which
confirms that the simulation achieves a reasonable level of asynchrony.

Evaluation item (d) measures how much bookkeeping is required to sustain active
users in the forums.
Even though the newbie rule~\code{4.a} in Table~\ref{fig.rules} is key to
combat Sybils, ideally it should not deny access to active users with low
reputation recurrently.
%
The evaluation counts the number of blocked messages requiring extra likes
after the welcoming likes (which are disconsidered) and calculates the ratio
over the total number of messages in the chain.
%This situation occurs when an user with low reputation tries to post multiple
%messages in a row without the necessary interval imposed by rule~\code{2}.
As an example, if $10$ users posted $110$ messages requiring $20$ likes, we
discount the $10$ initial messages and welcoming likes and find a ratio of
$10\%$ (\code{(20-10)/(110-10)}).
%
For the chat with $80$ users, we found a ratio of $3.7\%$.
For the newsgroup with $5000$ users, we found a ratio of $3.5\%$.
%
Considering that users were not aware of the reputation rules, the low ratios
indicate that their "natural" posting behavior matches the constraints of the
rules.
%As expected, the newsgroup has a lower bookkeeping given the higher interval
%between messages.
%
We assume that users would use the revoke mechanism to combat abusive content,
having no further effects on our evaluation.
%We also do not evaluate the quality of posts and how they would affect the
%reputation of users.

\begin{comment}
CHAT
timestamp user message
 798857 bytes

> DELL+ /x/tmp/fc-chat-20220529
- 3.7
- 455793 7419368 7875161
- erro (consensus ate 93\_*)
- 10491 / 1794 = 17%

> DELL+ /x/tmp/fc-chat-20220530
- 3.7
- 455793 7433845 7889638
- erro (consensus ate 93\_*)
- 10510 / 1851 = 18%

> DELL- /x/tmp/fc-chat-20220529
- 3.7
- 455793 7434379 7890172
- 125s, 50ms
- 10511 / 1852 = 18%

> DELL- /x/tmp/fc-chat-20220530
- 3.7
- 455793 7433590 7889383
- 150s, 50ms
- 10510 / 1850 = 18%

> NOTE 20220530
- 5.1
 455811
7577544
- 100ms / 130ms
- 10693 / 1811 = 17%

> NOTE 20220531
- 3.7
 455793
7434804
- 100ms / 130ms
- 10511 / 1851 = 18%

> NOTE 20220601
- 3.5
 455793
7384694
- 249s / 100ms
- 10456 / 1837 = 18%

-------------------------------------------------------------------------------

USENET
code From: Subject: Date: BODY
29.863.335 bytes

> DELL+ /x/tmp/fc-usenet-20220531
- 8.9
42.262.616
- 73s / 66ms
- 15659 / 2395

> DELL+ /x/tmp/fc-usenet-20220601
- 8.9
42357914
- 75s / 70ms
- 15682 / 2195

> DELL+ /x/tmp/fc-usenet-20220602
- 8.9
- 69s / 66ms
- 15669 / 2286

> DELL+ /x/tmp/fc-usenet-20220603
- 3.5
- 64s / 64ms
- 15416 / 2327

> DELL+ /x/tmp/fc-usenet-20220604
- 3.4
- 64s / 66ms
- 15419 / 2329

> DELL+ /x/tmp/fc-usenet-20220605
- 3.3
- 65s / 64ms
- 15391 / 2324

> DELL- /x/tmp/fc-usenet-20220601
- 8.7
42160398
- 52s / 70ms
- 15665 / 2198 = 14%

> DELL- /x/tmp/fc-usenet-20220602
- 60s / 70ms
- 15678 / 2091 = 13%

- lockdown strategy ok?

Discussion:
- do not scale
    - restart
- cannot evaluate behavior
    - dislike is immediate
- users are not aware of the policies
    -
- no images, videos
    - 128k enough for memes?

- Two cases
    - chat:
        - 2010-08-25 2012-11-01
        - 166277 messages
        - 13726470 bytes w/  metadata (date/time/user+msg)
        - 8592725 bytes w/o metadata (only msg)
        - 82/50 bytes/message (w/-w/o metadata)
        - 651 users
        - 227 messages per day

    - forum:
        - 1987-08-17 2012-05-28
        - 35366 messages
        - 117458214 bytes w/ metadata (date/time/user/*/+msg)
        - 3321 bytes/message
        - 13000 users
        - 3.8 messages per day

- number of forks
    - chat 30s per peer
    - e-mail 1h per peer
- time of consensus
\end{comment}

\section{Related Work}
\label{sec.related}

Many other systems have been proposed for decentralized content
sharing~\cite{p2p.survey,p2p.ecosystem}.
Here we consider the classes of publish-subscribe, federated, and peer-to-peer
protocols.

%\subsection{Publish-Subscribe Protocols}

Decentralized topic-based publish-subscribe protocols, such as
    \emph{XMPP}~\cite{pubsub.xmpp},
    \emph{ActivityPub}~\cite{pubsub.activitypub}, and
    \emph{gossipsub}~\cite{pubsub.gossipsub},
decouples publishers from subscribers in the network.
%
A key limitation of \emph{pubsubs} is that the brokers that mediate
communication still have a special role in the network, such as authenticating
and validating posts.
%
Nevertheless, some pubsubs do not rely on server roles, and instead, use
peer-to-peer gossip dissemination~\cite{pubsub.tera,pubsub.rappel,pubsub.stan,pubsub.vitis,pubsub.gossipsub,pubsub.rappel}.
Most of these protocols focus on techniques to achieve scalability and
performance, such as throughput, load balancing, and real-time relaying.
%
However, these techniques alone are not sufficient to operate permissionless
networks with malicious Sybils~\cite{pubsub.gossipsub2}.
Being generic protocols, pubsubs are typically unaware of the applications
built on top of them.
%
In contrast, as stated in Section~\ref{sec.freechains}, the pubsub of \FC is
conceptually at the application level and is integrated with the semantics of
chains, which already verifies blocks at publishing time.
For instance, to flood the network with posts, malicious peers need to spend
reputation, which takes hours to recharge (rule~\code{2} in
Table~\ref{fig.rules}).
In addition, blocked posts (Figure~\ref{fig.state}) are not a concern either,
because they have limited reachability.
Another advantage of a tighter integration between the application and protocol
is that Merkle~DAGs simplify synchronization, provide persistence, and prevent
duplication of messages.
%
In summary, \FC and \emph{pubsub} middlewares operate at different network
layers, which suggests that \FC could benefit of the latter to manage the
interconnections between peers.

%\subsection{Federated Protocols}

Federated protocols, such as e-mail, allow users from one domain to seamlessly
exchange messages with users of other domains.
\emph{Diaspora}, \emph{Matrix}, and \emph{Mastodon} are recent federations for
social media, chat, and microblogging~\cite{p2p.ecosystem}, respectively.
%
As a drawback, identities in federations are not portable across domains, which
may become a problem when servers shutdown or users become unsatisfied with the
service~\cite{fed.distributed}.
In any of these cases, users have to grab their content, move to another
server, and announce a new identity to followers.
%
Moderation is also a major concern in federations~\cite{p2p.ecosystem}.
As an example, messages crossing domain boundaries may be subject to different
policies that might affect delivery.
With no coordinated consensus, it is difficult to make pervasive public forums
practical.
%
For this reason, Matrix supports a permissioned moderation system%
\footnote{Matrix moderation: \url{https://matrix.org/docs/guides/moderation}},
but which applies only within clients, after the messages have already been
flooded in the network.
%
As a counterpoint, federated protocols seem to be more appropriate for
real-time applications such as large chats rooms.
The number of hops and header overhead can be much smaller in client-server
architectures compared to peer-to-peer systems, which typically include message
signing, hash linking, and extra verification rules.

%\subsection{Peer-to-Peer Protocols}

Regarding peer-to-peer protocols, Bitcoin~\cite{p2p.bitcoin} is probably the
most successful permissionless network, but serves specifically for electronic
cash.
IPFS~\cite{p2p.ipfs} and Dat~\cite{p2p.dat} are data-centric protocols for
hosting large files and applications, respectively.
Scuttlebutt~\cite{p2p.scuttlebutt} and Aether~\cite{p2p.ecosystem} are closer
to \FC goals and cover human-centric \Xon and \Xnn public communication,
respectively~\cite{fcs.sbseg20}.

Bitcoin adopts proof-of-work to achieve consensus, which does not solve the
centralization issue entirely, given the high costs of equipment and energy.
Proof-of-stake is a prominent alternative~\cite{p2p.proofs} that acknowledges
that centralization is inevitable, and thus uses a function of time and wealth
to elect peers to mint new blocks.
As an advantage, these proof mechanisms are generic and apply to multiple
domains, since they depend on an extrinsic scarce resource.
% (i.e., the richer gets richer)
In contrast, we chose an intrinsic resource, which is authored content in the
chains themselves.
We believe that human work grows more linearly with effort and is not directly
portable across chains with different topics.
These hypotheses support the intended decentralization of our system.
%
Another distinction is that generic public ledgers require permanent
connectivity to avoid forks, which opposes our local-first principle.
This is because a token transaction only has value as part of the longest
chain.
This is not the case for a local message exchange between friends, which has
value in itself.

IPFS~\cite{p2p.ipfs} is centered around immutable content-addressed data, while
Dat~\cite{p2p.dat} around mutable pubkey-addressed data.
IPFS is more suitable to share large and stable content such as movies and
archives, while Dat is more suitable for dynamic content such as web apps.
%
Both IPFS and Dat use DHTs as their underlying architectures, which are optimal
to serve large and popular content, but not for search and discovery.
In both cases, users need to know in advance what they want, such as the exact
link to a movie or a particular identity in the network.
%
On the one hand, DHTs are probably not the best architecture to model
decentralized human communication with continuous feed updates.
On the other hand, replicating large files across the network in Merkle~DAGs is
also impractical.
An alternative is to use DHT links in Merkle payloads to benefit from both
architectures.

Scuttlebutt~\cite{p2p.scuttlebutt} is designed around public identities that
follow each other to form a graph of connections.
This graph is replicated in the network topology as well as in data storage.
For instance, if identity $A$ follows identity $B$, it means that the computer
of $A$ connects to $B$'s in a few hops and also that it stores all of his posts
locally.
Scuttlebutt is aligned to \Xon public identity chains of \FC.
%
For group \Xnn communication, Scuttlebutt uses the concept of channels, which
are in fact nothing more than hash tags (e.g. \emph{\#sports}).
Authors can tag posts, which appear not only in their feeds but also in local
virtual feeds representing these channels.
However, users only see channel posts from authors they already follow.
In practice, channels simply merge friends posts and filter them by tags.
In theory, to read all posts of a channel, a user would need to follow all
users in the network (which also implies storing their feeds).
A limitation of this model is that new users struggle to integrate in channel
communities because their posts have no visibility at all.
As a counterpoint, channels are safe places that do not suffer from abuse.
%
\begin{comment}
For group \Xnn communication, \FC relies on the reputation system to prevent
abuse, since new users require at least one like for visibility in the
community.
Also, a forum chain stores its own posts only, and are not simply tags over the
entire content in the network.
\end{comment}

Aether~\cite{p2p.ecosystem} provides peer-to-peer public communities aligned
with \Xnn public forums of \FC.
A fundamental difference is that Aether is designed for ephemeral, mutable
posts with no intention to enforce global consensus across peers.
Aether employs a very pragmatic approach to mitigate abuse in forums.
It uses established techniques, such as proof-of-work to combat SPAM, and an
innovative voting system to moderate forums, but which affects local instances
only.
In contrast, \FC relies on its permissionless reputation and consensus
mechanisms for moderation.

\section{Conclusion}
\label{sec.conclusion}

In this paper, we propose a permissionless consensus and reputation mechanism
for social content sharing in peer-to-peer networks.
We enumerate four main contributions:
    (i)   human authored content as a scarce resource
          (\emph{proof-of-authoring});
    (ii)  diversified public forums, each as an independent blockchain with
          subjective moderation rules; and
    (iii) abusive content removal preserving data integrity.

The key insight of the consensus mechanism is to use the human authoring
ability as a scarce resource to determine consensus.
This contrasts with extrinsic resources, such as CPU power, which are
dispendious and not evenly distributed among people.
%
Consensus is backed by a reputation system in which users can rate posts with
likes and dislikes, which transfer reputation between them.
The only way to forge reputation is by authoring new content under the
judgement of other users.
This way, reputation generation is expensive, while verification is cheap and
decentralized.

The reputation and consensus mechanism is integrated into \FC, a peer-to-peer
protocol that offers multiple arrangements of public and private
communications.
Users have the power to create public forums of interest and apply diverse
moderation policies.
In particular, users can revoke content considered abusive according to the
majority, not depending on centralized authorities.
%
We simulate the behavior of existing chat and newsgroup forums as if they were
using \FC to show the practicability of the protocol as a descentralized
alternative for public forums.

Finally, we do not claim that the proposed reputation system enforces ``good''
human behavior in any way.
Instead, it provides a transparent and quantitative mechanism to help users
understand the evolution of forums and act accordingly.
Human creativity contrasts with plain economic resources (e.g.,
proof-of-work), which do not appraise social interactions and also tend
to concentrate over the time.
%
\begin{comment}

where else can we use this?
    - reddit

- Users can fork or recreate the chain.
  Unlike bitcoin, the value is not on the size, but the cohesion is users.
- Safe for niche topics and minority groups
- Fulfill the expectations of part of the community
- Part from the principle that the pioneers want it to decentralize, otherwise would create a public identity chain fig2

- INSIGHT
 Uso descentralizado
 Emissão descentralizada e restrita
 Criação difícil / Verificação fácil

- users that work twice, bots

- summary
    - parallel w/ bitcoin
        - ponte entre reputacao -> sybil -> consenso
    - no problem w/ forks
    We reach a similar solution to Bitcoin but adapted to our domain
    - we need consensus to make the reputation system work
    - we need reputation system to reach consensus
    - same virutous cycle as bitcoin
    - The more CPU work is done, the stronger becomes the proposal, the more peers
        follow it, the more tokens are mined.
        There is a strong association between work, profit and consensus that enables
        Bitcoin as a peer-to-peer cash system.
    - reputation system to rate messages
        - positively (to distinguish from excess), or
        - negatively (to block SPAM, fake news, illegal content)
    - some sort of scarcity (work)
        - you like you loose
        - you work you get
        - otherwise sybil, "likes" abuse
        - incentives for continuous, good quality posts
    - Like \emph{bitcoins}, \reps are scarce, hard to generate, and easy to verify.
      Unlike \emph{bitcoins}, \reps .

\end{comment}

\begin{comment}
- content
- CPU work to create blocks, easy objective verification, 50+1 attack, collapse
- Human work to create content, easy subjective verification, 50+1 attack,
  community fork (actually encouraged)
- unique identity based on CPU
- here based on post quality
- not CDN: Content delivery network

(b) double spend of coins/reps is solved by total ordering all
Users can like \& dislike posts, which transfer reputation between them.
Reputation is created from news

Just like Bitcoin reaches consensus with the longest chain
uses mining to

(excess, SPAM, fake, abuse, illegal)
\end{comment}

\bibliographystyle{abbrv}
\bibliography{freechains}

\end{document}

### 2023-05-10 freechains/freechains.tex
...

@@
 %
-Consensus is key to eradicate Sybil attacks~\cite{p2p.sybil}, which are the
-major threat to decentralized applications in general: without consensus, it
-is not possible, \emph{at the protocol level}, to distinguish between correct
-and malicious users in order to satisfy item (iii).
+Consensus is key to distinguish, \emph{at the protocol level}, between correct
+and malicious users, and therefore, satisfy item (iii).
+Without consensus, each client needs local anti-abuse policies, such as SPAM
+filters and blacklists, but which only apply after the fact, when the abusive
+messages have already been flooded in the network.
+%
 %fundamental in collaborative applications, such as open discussion
@@
           attacks;
-    (ii)  they lean towards concentration of power due to scaling effects; and
-    (iii) they impose an external economic cost to use the protocol.
-These issues threaten our original decentralization goals.
-In particular, a unique timeline implies that all Internet content should be
-subject to the same consensus rules, which neglects all subjectivity that is
-inherent to social content.
+    (ii)  they lean towards concentration of power due to scaling effects;
+    (iii) they impose an external economic cost to use the protocol; and
+    (iv)  they rely exclusively on objective rules to reach consensus.
+%
+These issues threaten our original decentralization goals, and more
+importantly, they renounce subjectivity, which is inherent to content
+moderation.
+%
+For instance, a unique timeline implies that all Internet content should be
+subject to the same consensus rules, while objective rules alone cannot
+distinguish abuse amid content.
 %
@@
 illegal content (e.g., hate speech).
-
-In this work, we adapt the idea of scarce resources to reach consensus, but in
-the context of social content sharing.
-%
-Our first contribution is to recognize the actual published contents as the
-protocol scarce resources, since they require human work.
+For instance, Bitcoin transactions permanently store arbitrary human-authored
+content~\cite{btc.data}, including harmful texts and images~\cite{btc.content}.
+
+In this work, we adapt the idea of scarce resources to reach consensus and
+eradicate Sybils, but in the context of social content sharing.
+%
+Our first contribution is to recognize the actual published contents themselves
+as the scarce resources, since they require human work.
 %
@@
 %
-Due to decentralization, posts in a timeline form a causal graph with only
-partial order, which we promote to a total order based on the reputation of
-authors.
+Due to decentralization and network partitions, posts in a timeline form a
+causal graph with only partial order, which we promote to a total order based
+on the reputation of authors.
 The consensus order is fundamental to detect conflicting operations, such as
@@
 %
-Our second contribution is to allow that users create diversified forums of
+Our second contribution is to allow users to create diversified forums of
 interest (instead of a singleton blockchain), each counting as an independent
@@
 Users have the power to revoke posts with dislikes, and peers are forced to
-remove payloads, only forwarding associated metadata.
+remove payloads, only forwarding associated blockchain metadata.
 %
@@
 
-\newpage
 \section{The Reputation and Consensus Mechanism}
@@
 % and make peer-to-peer public forums practical.
+
+\begin{table}
+\centering
+\includegraphics[width=0.75\textwidth]{general.png}
+\caption{General reputation operations in public forums.}
+\label{fig.general}
+\end{table}
 
@@
 \footnote{
-    Note that to some extent work from AI may be indistinguishable from humans,
-    but its frequency and relevance are still subject to human evaluation,
-    which is subjective and evolves over time.
+    Note that to some extent AI content may be indistinguishable from humans,
+    but its frequency and relevance in forums are still subject to human
+    evaluation in our proposal.
 }
-
-\begin{table}
-\centering
-\includegraphics[width=0.75\textwidth]{general.png}
-\caption{General reputation operations in public forums.}
-\label{fig.general}
-\end{table}
 
@@
 %
-Instead, considering the context of public forums, we can take advantage of the
-human authoring ability as an intrinsic resource.
+Instead, considering the context of public forums, we propose to take advantage
+of the human authoring ability as an intrinsic resource.
 Creating new content is hard and takes time, but is comparatively easy to
@@
 %Additionally,
-%
-Nevertheless, scarce operations are not yet sufficient because they demand
-consensus to establish an order in time across the network to prevent
-inconsistent operations.
+
+Nevertheless, posts scarcity is not yet sufficient because they still demand a
+consensual order in time across the network to prevent inconsistent operations.
 As an example, consider a malicious author with a single unit of \reps posting
-new messages using multiple peers at the same time.
+SPAM messages from multiple peers at the same time.
 According to the \emph{Expense} rule of Table~\ref{fig.general}, only one of
-these messages should be accepted.
+these messages should be accepted by the network.
 However, without consensus, it is not possible to globally determine which
 message to accept, since each peer would supposedly accept the first message it
-sees.
+receives.
 On the one hand, in order to validate operations consistently, we need the same
 message ordering across all peers in the network.
-On the other hand, we typically can only represent peer-to-peer public forums
-as DAGs with causal relationships between messages, which provides at most
-partial order.
+On the other hand, due to network partitions, we can only represent public
+forums as DAGs with causal relationships between messages, which provides at
+most partial order.
+
+\subsection{Basic Reputation Consensus}
 
@@
 \centering
-\includegraphics[width=0.9\textwidth]{reps2.png}
+\includegraphics[width=0.9\textwidth]{reps.png}
 \caption{
-    (A) A public forum DAG with a common prefix and two branches.
-    (B) Total order between blocks of the DAG after consensus.
+    (A) A public forum DAG with a common prefix, two branches, and a common sufix.
+    (B) Total order between posts of the DAG after consensus.
 }
@@
 
-Therefore, our key idea to stablish consensus in forum DAGs is to favor forks
-with posts from users that constitute the majority of the reputation in the
-network.
+Our key idea to stablish consensus in forum DAGs is to favor forks with posts
+from users that constitute the majority of the reputation in the network.
 These forks have more associated work from active users and are analogous to
@@
 
-Figure~\ref{fig.reps}.A illustrates the consensus criteria.
-%, still abstractly, since we did not discuss the actual rules for content creation and rating.
-A public forum DAG has a common prefix with signed posts from users $a$, $b$,
-and $c$.
-Let's assume that within the prefix, users $a$ and $b$ have contributed with
-better content and have more reputation combined than $c$ has alone (i.e.,
-$8+5>3$).
-%
-After the prefix, the forum forks in two branches:
-in \code{branch-1}, only user $c$ remains active and we see that new users $x$
-and $y$ (with no previous reputation in the common prefix) generate a lot of
-new content;
-in \code{branch-2}, only users $a$ and $b$ participate but with less activity.
-Nevertheless, \code{branch-2} would be ordered first in this case because,
-before the forking point, $a$ and $b$ have more reputation than $c$, $x$, and
-$y$ combined.
-%
-User $c$ here might represent a malicious user trying to cultivate fake
-identities $x$ and $y$ in separate of the network during weeks to accumulate
-\reps.
-%However, the whole malicious \code{branch-1} is vulnerable because users in
-%\code{branch-2} with more previous reputation take the priority and can
-%overthrow user $c$.
-
-Figure~\ref{fig.reps}.B indicates the resulting consensus order between blocks
-in the forum.
-All operations in \code{branch-2} appear before any operation in
-\code{branch-1}.
-%
+Figure~\ref{fig.reps}.A illustrates the evolution of a forum DAG in the
+presence of a network partition:
+A common prefix has signed posts from users $a$, $b$, and $c$, when they were
+still all connected.
+Then, user $c$ disconnects from $a$ and $b$, and evolves in separate along with
+new users $x$ and $y$ with posts in \code{branch-1}.
+In the meantime, $a$ and $b$ participate with posts in \code{branch-2}, but
+with less activity.
+After some time, the partitions reconnect and new posts from $x$, $b$ and $c$
+merge the branches with a common suffix.
+
+Even in the ordinary case of a legit temporary network partition, note that the
+two branches may still show conflicts when merging.
+For instance, posts with dislike operations in one branch could possibly block
+posts from users in the other branch.
+Therefore, we need to order the branches with a global consensus, such that we
+can validate operations consistently across the network.
+%
+Note also that, to reach consensus, we should preferably rely on the maximum
+information that all branches agree on, which is exactly the posts in the
+common prefix.
+
+In Figure~\ref{fig.reps}.A, we assume that within the common prefix, users $a$
+and $b$ have contributed with better content and therefore have more reputation
+combined than $c$ has alone (i.e., $8+5>3$).
+%
+Based on that information, Figure~\ref{fig.reps}.B indicates the resulting
+consensus order between posts in the forum DAG:
+    the common prefix comes first,
+    followed by all posts in \code{branch-2},
+    followed by all posts in \code{branch-1},
+    followed by the common suffix.
+%
+The \code{branch-2} with users $a$ and $b$ take priority over \code{branch-1}
+with user $c$ because, before the forking point, $a$ and $b$ have more
+reputation than $c$, $x$, and $y$ combined.
+
 At any point in the consensus timeline, if an operation fails, all remaining
-blocks in the offending branch are removed from the primary DAG.
+posts in the offending branch are removed from the primary DAG.
 As an example, suppose that the last post by $a$ (in gray) is a dislike to user
@@
 Then, it's possible that the last post by $c$ (in red), now (suppose) with
-\nreps{0}, is rejected together with all posts by $y$ and $x$ in sequence.
-%
-The removal rule for the remaining blocks is a direct consequence of tamper
-proof Merkle~DAGs~\cite{p2p.ipfs,p2p.bitcoin} used in permissionless protocols,
-which cannot support graph modifications.
-
-Note that the consensus order exists only for accountability purposes, and is
-a view of the primary DAG structure.
-%
-Note also that users in the branch with more reputation can react to attacks
-even after the fact.
-For instance, users $a$ and $b$ can pretend that they did not yet see
-\code{branch-1} and post extra dislikes to user $c$ from \code{branch-2} so
-that a further merge removes all blocks of \code{branch-1} from the DAG.
-%
+\nreps{0}, is rejected together with all posts in sequence, including those in
+the common suffix.
+%
+This rejection rule for remaining posts is a direct consequence of tamper proof
+Merkle~DAGs~\cite{p2p.ipfs,p2p.bitcoin} typically used in permissionless
+protocols, which cannot support graph modifications.
+
+The basic consensus rule is therefore straightforward:
+Whenever a fork is found, we create sets of users with posts in each branch
+($B1=\{c,x,y\}$ and $B2=\{a,b\}$).
+Then, we take each set, and sum its users \reps based on the common prefix
+($S1=3$ and $S2=13$), since it is the maximum information that all branches
+agree.
+Finally, we order the branches from the highest sum of \reps found
+(Figure~\ref{fig.reps}.B).
+While applying the branches in order, if any post operation fails, all
+remaining posts are rejected, as if they never existed.
+
 There are some other relevant considerations about forks and merges:
@@
 Peers that first received branches with less reputation will need to reorder
-all blocks starting at the forking point.
+all posts starting at the forking point.
 This might involve removing content in the end-user software.
@@
 %and do not need to recompute anything.
-This behavior is expected to happen in the majority of the network.
+This behavior is expected to happen in the majority of the connected network.
 %
@@
 However, the longer a peer remains disconnected, the more conflicting
-operations it may see, and the higher are the chances of rejection when
+operations it may see, and the higher are the chances of reordering when
 rejoining.
 
-As a counterpoint to the consensus order in Figure~\ref{fig.reps}.B, maybe
-users $a$ and $b$ have abandoned the forum for months, and thus \code{branch-1}
-is actually legit.
+In summary, it is important to remark that
+    (i) branches always introduce conflicts, which
+    (ii) require consensus to order operations, which
+    (iii) is based on reputation of an agreed common prefix,
+    (iv) leading to a global and deterministic and result.
+%
+Note that the consensus order exists only for accountability purposes, and is
+a view of the DAG structure, which remains as the primary source of data.
+
+\subsection{Malicious Behaviors and Hard Forks}
+
+Now, consider a malicious scenario in which user $c$ in Figure~\ref{fig.reps}
+intentionally disconnected from the network for weeks to cultivate fake
+identities $x$ and $y$ and accumulate \reps.
+When merging, these fake users become legit and now constitute the majority of
+the forum in terms of reputation.
+%
+However, because the common prefix is the basis of consensus, users in the
+branch with more reputation can still react even after the attack.
+For instance, users $a$ and $b$ can pretend that they did not yet see
+\code{branch-1} and post extra dislikes to user $c$ from \code{branch-2} before
+the common suffix so that a further merge removes all posts of \code{branch-1}
+from the DAG.
+%
+The example shows that this kind of attack is innocuous because it requires at
+least half of the \reps accumulated in the common prefix.
+%
+The same countermeasure can be used against a minority of users colluding to
+boost or protect their own posts.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{n100-d7.png}
+\caption{
+    Stable consensus permanently freezes the order of posts once they cross
+    the threshold.
+    The unstable order may still be affected by incoming branches.
+}
+\label{fig.hard}
+\end{figure}
+
+As a counterpoint, maybe users $a$ and $b$ have abandoned the forum for months,
+and thus \code{branch-1} is actually legit.
 In this case, users $a$ and $b$ might be the ones trying to take over the
@@
 for a long period of time.
-In any case, it is unacceptable that a very old remote branch affects a long
-active local forum.
+In any case, it is unacceptable for users that a very old remote branch affects
+a long active forum.
 %
 For this reason, the consensus algorithm includes an extra constraint that
-prevents long-lasting local branches to merge, creating \emph{hard forks} in
-the network.
+prevents long-lasting local branches to merge, which leads to \emph{hard forks}
+in the network.
 A hard fork occurs when a local branch crosses a predetermined and irreversible
@@
 never synchronize again.
-More than simply numeric disputes, hard forks represent social conflicts in
+More than simple numeric disputes, a hard fork represents a social conflict in
 which reconciling branches is no longer possible.
@@
 Figure~\ref{fig.hard} illustrates hard forks by distinguishing \emph{stable
-consensus}, which cannot be reordered, from \emph{unstable consensus}, which may
-still be affected by incoming branches.
-The activity threshold counts backwards, from the latest local block in the
-unstable consensus.
-
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{n100-d7.png}
-\caption{
-    Stable consensus freezes the order of blocks once they cross the threshold.
-    The unstable order may still be affected by incoming branches.
-}
-\label{fig.hard}
-\end{figure}
+consensus}, which cannot be reordered, from \emph{unstable consensus}, which
+may still be affected by incoming branches.
+The activity threshold counts backwards, starting from the newest local post
+in the unstable consensus to older posts until they are permanently frozen.
 
@@
     \item otherwise, branches are ordered by an arbitrary criteria, such as
-          lexicographical order of the block hashes immediately after the
+          lexicographical order of the post hashes immediately after the
           common prefix.
@@
 
+\subsection{Content Removal}
+
 Finally, we consider that revoking posts is fundamental in the context of
@@
 However, if the number of dislikes exceeds an arbitrary threshold (e.g., more
-dislike than likes), the block becomes \code{REVOKED} and its payload is not
+dislike than likes), the post becomes \code{REVOKED} and its payload is not
 retransmitted to other peers.
-Note that a block hash does not depend on its associated payload, but only on
+Note that a post hash does not depend on its associated payload, but only on
 the payload hash.
@@
 somewhere and peers can request it when synchronizing again.
+
+\subsection{Byzantine Peers}
+
+Since posts are signed and structured in Merkle~DAGs, ~\cite{p2p.ipfs,p2p.bitcoin} used in permissionless
+Regarding Byzantine-faulty peers, which deviate from the protocol
+p2p.byz
+When a node deviates from the specified protocol, we
+call it Byzantine-faulty (or simply Byzantine), regardless of
+whether the deviation is by accident or by malice [25]. Byzan-
+tine behaviour is not always detectable by other nodes, since
+a Byzantine node may attempt to hide its protocol violations.
+
+To conclude this section, we should keep in mind that our consensus algorithm
+is based on subjective evaluations of users and posts.
+For this reason, unlike Bitcoin, in which forks are discouraged and a single
+blockchain must prevail, public forums can split and diverge indiscriminately.
+For instance, users can explicitly apply hard forks at any point in time or
+``reboot'' a community from a previous state.
+Ultimately, the consensus algorithm provides a transparent mechanism to help
+users understand the evolution of public forums and act accordingly.
 
@@
 Note that chains with the same name but different pioneers would be
-incompatible because the hash of genesis blocks depend on the pioneers' public
+incompatible because the hash of genesis posts depend on the pioneers' public
 keys.
@@
 Evaluation item (a) measures the protocol overhead due to blockchain metadata,
-which consists of a timestamp, author signature, and hashes (block id, payload,
+which consists of a timestamp, author signature, and hashes (post id, payload,
 backlinks, and likes).
@@
 chain.
-We measured the time to sort blocks in a local DAG both for the first time
+We measured the time to sort posts in a local DAG both for the first time
 (without any caches), and incrementally (from stable consensus caches).
@@
 conceptually at the application level and is integrated with the semantics of
-chains, which already verifies blocks at publishing time.
+chains, which already verifies posts at publishing time.
 For instance, to flood the network with posts, malicious peers need to spend

### 2023-05-11 freechains/freechains.tex
...

@@
 
+\begin{comment}
 \begin{resumo} 
@@
 \end{resumo}
+\end{comment}
 
@@
 \emph{consensus}, from now on) in the context of decentralized content sharing.
+
+In this sense, consensus 
+eradicate Sybil, which are the
+major threat to decentralized applications in general.
+The reason
+Furthermore, without consensus, it is not possible,
 \end{comment}
@@
 Bitcoin~\cite{p2p.bitcoin} is the first permissionless protocol to resist
-Sybils through consensus.
+Sybil attacks~\cite{p2p.sybil}.
 Its key insight is to rely on a scarce resource---the \emph{proof-of-work}---to
@@
 
+To conclude this section, we should keep in mind that our consensus algorithm
+is based on subjective evaluations of users and posts.
+For this reason, unlike Bitcoin, in which forks are discouraged and a single
+blockchain must prevail, public forums can split and diverge indiscriminately.
+For instance, users can explicitly apply hard forks at any point in time or
+``reboot'' a community from a previous state.
+Ultimately, the consensus algorithm provides a transparent mechanism to help
+users understand the evolution of public forums and act accordingly.
+
+\subsection{Content Removal}
+
 \begin{figure}
@@
 
-\subsection{Content Removal}
-
 Finally, we consider that revoking posts is fundamental in the context of
@@
 
-\subsection{Byzantine Peers}
-
-Since posts are signed and structured in Merkle~DAGs, ~\cite{p2p.ipfs,p2p.bitcoin} used in permissionless
-Regarding Byzantine-faulty peers, which deviate from the protocol
-p2p.byz
-When a node deviates from the specified protocol, we
-call it Byzantine-faulty (or simply Byzantine), regardless of
-whether the deviation is by accident or by malice [25]. Byzan-
-tine behaviour is not always detectable by other nodes, since
-a Byzantine node may attempt to hide its protocol violations.
-
-To conclude this section, we should keep in mind that our consensus algorithm
-is based on subjective evaluations of users and posts.
-For this reason, unlike Bitcoin, in which forks are discouraged and a single
-blockchain must prevail, public forums can split and diverge indiscriminately.
-For instance, users can explicitly apply hard forks at any point in time or
-``reboot'' a community from a previous state.
-Ultimately, the consensus algorithm provides a transparent mechanism to help
-users understand the evolution of public forums and act accordingly.
+\subsection{Synchronization and Byzantine Peers}
+
+Similarly to Bitcoin, the underlying network protocol replicates on each peer
+the full state of the Merkle-DAG representing the causal relationships between
+the messages.
+When peers synchronize, they mutually synchronize missing branches such that
+they reach the same DAG structure~\cite{p2p.byz}.
+As they synchronize, each peer runs the consensus algorithm locally and
+verifies if all received posts are consistent.
+
+Because the protocol relies on a tamper-proof DAGs, and because each peer
+self-verifies the posts as they synchronize, the protocol becomes tolerant to
+Byzantine-faulty peers~\cite{lamport.byz}.
+%
+No information coming from any peer is trusted a priori.
+%
+For instance, a malicious peer that tries to synchronize branches with erratic
+data can be detected on the first inconsistent post.
+This allows the correct peer to close the connection immediately and blacklist
+the malicious peer from future connections.
+
+Therefore, although Byzantine peers may still perform some kind of denial of
+service attacks in the network, they cannot modify the DAG structure, create
+new content (without performing work), or impede that two correct nodes
+synchronize directly.
 
@@
 
-{\footnotesize
+{\scriptsize
 \begin{verbatim}
@@
 chains.
-Fully peer-to-peer systems cannot rely on logins or {\footnotesize CAPTCHAs}
+Fully peer-to-peer systems cannot rely on logins or {\scriptsize CAPTCHAs}
 due to the lack of a central authority.

### 2023-05-11 freechains/freechains.tex
...

@@
 \title{
-    Peer-to-Peer Permissionless Consensus via Authoring Reputation
+    Peer-to-Peer Permissionless Consensus via Reputation
 }
@@
 and robust access.
-On the other hand, they concentrate power by collecting users' data and
-``algorithmizing'' consumption.
+On the other hand, they concentrate too much power over its users, holding and
+monetizing over their private data and behavior.
 %
 Peer-to-peer alternatives~\cite{p2p.survey} eliminate intermediaries, but
-strive to achieve consistency while dealing with malicious users.
+strive to achieve consistency when dealing with malicious users.
 %given the decentralization of authority and infrastructure.
@@
 %
-Consensus is key to distinguish, \emph{at the protocol level}, between correct
-and malicious users, and therefore, satisfy item (iii).
-Without consensus, each client needs local anti-abuse policies, such as SPAM
-filters and blacklists, but which only apply after the fact, when the abusive
-messages have already been flooded in the network.
+Consensus is key to distinguish malicious behavior \emph{at the protocol
+level}, and therefore, satisfy item (iii).
+Without consensus, each client must rely on local anti-abuse policies, such as
+SPAM filters and blacklists, but which only apply a posteriori, when the
+abusive messages have already been flooded in the network.
 %
@@
 %Otherwise, they become impractical due to excess and abuse (e.g., SPAM and
-%illegal content).
+%illicit content).
 %At the very core, consensus is
@@
 content in the middle of a blockchain, which is inadmissible considering
-illegal content (e.g., hate speech).
+illicit content (e.g., hate speech).
 For instance, Bitcoin transactions permanently store arbitrary human-authored
@@
 %
-Our first contribution is to recognize the actual published contents themselves
-as the scarce resources, since they require human work.
+Our first contribution is to recognize the actual published contents as the
+scarce resources themselves, since they require human work.
 %
 Work is manifested as new posts, which if approved by others, reward authors
-with reputation tokens, which are used to evaluate other posts with likes and
-dislikes.
+with reputation tokens, which are further used to evaluate other posts with
+likes and dislikes.
 With such \emph{proof-of-authoring} mechanism, token generation is expensive,
@@
 In Section~\ref{sec.design}, we describe the design of the reputation and
-consensus mechanism for public forums.
+consensus mechanism for public forums and discuss some security threats.
 In Section~\ref{sec.freechains}, we describe the concrete reputation rules we
-implemented for \FC.
-In Section~\ref{sec.evaluation}, we evaluate the performance of the protocol
-in real-world public forums.
+implemented for \FC and evaluate the performance of the protocol in real-world
+public forums.
 In Section~\ref{sec.related}, we compare our system with publish-subscribe
@@
 system.
-The only way to generate new \reps is to post content that other users approve,
-which demands non-trivial work resistant to automation.%
+The only way to generate new \reps is to post content that other users
+tolerate, which demands non-trivial work resistant to automation.%
 \footnote{
@@
 
-Nevertheless, posts scarcity is not yet sufficient because they still demand a
-consensual order in time across the network to prevent inconsistent operations.
-As an example, consider a malicious author with a single unit of \reps posting
+Nevertheless, posts scarcity is not yet sufficient to combat Sybils because
+consensual order is still required to prevent inconsistent operations.
+For instance, consider a malicious author with a single unit of \reps posting
 SPAM messages from multiple peers at the same time.
@@
 On the other hand, due to network partitions, we can only represent public
-forums as DAGs with causal relationships between messages, which provides at
-most partial order.
+forums as DAGs (directed acyclic graphs) with causal relationships between
+messages, which provides at most partial order.
 
@@
     (ii) measuring the quality of posts, and
-    (iii) revoking abusive posts (SPAM, fake news, illegal content, etc).
+    (iii) revoking abusive posts (SPAM, fake news, illicit content, etc).
 %
@@
         - positively (to distinguish from excess), or
-        - negatively (to block SPAM, fake news, illegal content)
+        - negatively (to block SPAM, fake news, illicit content)
     - some sort of scarcity (work)
@@
 
-(excess, SPAM, fake, abuse, illegal)
+(excess, SPAM, fake, abuse, illicit)
 \end{comment}

### 2023-05-12 freechains/freechains.tex
...

@@
 \includegraphics[width=0.75\textwidth]{general.png}
-\caption{General reputation operations in public forums.}
+\caption{General reputation rules in public forums.}
 \label{fig.general}
@@
 threshold.
-Table~\ref{fig.general} summarizes the reputation operations and their goals.
-To prevent Sybils, users with no \reps cannot perform these operations,
+Table~\ref{fig.general} summarizes the general reputation rules and their
+goals.
+To prevent Sybils, users with no \reps cannot operate under these rules,
 requiring a welcoming like from any other user already in the system.
@@
 \footnote{
-    Note that to some extent AI content may be indistinguishable from humans,
-    but its frequency and relevance in forums are still subject to human
-    evaluation in our proposal.
+    Note that to some extent AI content may be indistinguishable from human
+    content, but its frequency and relevance in forums is still subject to
+    human evaluation in our proposal.
 }
@@
 \subsection{Basic Reputation Consensus}
+\label{sec.design.basic}
 
@@
 %
-In technical terms, we can adapt a topological sorting algorithm to favor
+In technical terms, we adapt a topological sorting algorithm to favor
 reputation when deciding between branches in a forum DAG.
@@
 still all connected.
+%
+We assume that within the common prefix, users $a$ and $b$ have contributed
+with better content and therefore have more reputation combined than $c$ has
+alone (i.e., $8+5>3$).
+We will use this information when ordering branches in the consensus algorithm.
+%
 Then, user $c$ disconnects from $a$ and $b$, and evolves in separate along with
 new users $x$ and $y$ with posts in \code{branch-1}.
-In the meantime, $a$ and $b$ participate with posts in \code{branch-2}, but
-with less activity.
+In the meantime, $a$ and $b$ participate with posts in \code{branch-2}.
 After some time, the partitions reconnect and new posts from $x$, $b$ and $c$
@@
 
-Even in the ordinary case of a legit temporary network partition, note that the
-two branches may still show conflicts when merging.
-For instance, posts with dislike operations in one branch could possibly block
-posts from users in the other branch.
+Note that even non-malicious network partitions may experience conflicts when
+merging branches.
+For instance, dislike operations in one branch to common past posts could
+possibly block posts from users in the other branch, depending on how they are
+ordered.
 Therefore, we need to order the branches with a global consensus, such that we
@@
 
-In Figure~\ref{fig.reps}.A, we assume that within the common prefix, users $a$
-and $b$ have contributed with better content and therefore have more reputation
-combined than $c$ has alone (i.e., $8+5>3$).
-%
-Based on that information, Figure~\ref{fig.reps}.B indicates the resulting
-consensus order between posts in the forum DAG:
+Therefore, based on the reputation of users in the common prefix,
+Figure~\ref{fig.reps}.B indicates consensus order between posts in the forum
+DAG:
     the common prefix comes first,
@@
 %
-The \code{branch-2} with users $a$ and $b$ take priority over \code{branch-1}
+The \code{branch-2} with users $a$ and $b$ takes priority over \code{branch-1}
 with user $c$ because, before the forking point, $a$ and $b$ have more
 reputation than $c$, $x$, and $y$ combined.
-
-At any point in the consensus timeline, if an operation fails, all remaining
-posts in the offending branch are removed from the primary DAG.
-As an example, suppose that the last post by $a$ (in gray) is a dislike to user
-$c$.
-Then, it's possible that the last post by $c$ (in red), now (suppose) with
-\nreps{0}, is rejected together with all posts in sequence, including those in
-the common suffix.
-%
-This rejection rule for remaining posts is a direct consequence of tamper proof
-Merkle~DAGs~\cite{p2p.ipfs,p2p.bitcoin} typically used in permissionless
-protocols, which cannot support graph modifications.
 
@@
 agree.
-Finally, we order the branches from the highest sum of \reps found
+Finally, we order the branches based on the highest sum of \reps found
 (Figure~\ref{fig.reps}.B).
+
 While applying the branches in order, if any post operation fails, all
-remaining posts are rejected, as if they never existed.
+remaining posts are rejected and removed from the DAG, as if they never
+existed.
+As an example, suppose that the last post by $a$ (in gray) is a dislike to user
+$c$.
+Then, it's possible that the last post by $c$ (in red), now suppose with
+\nreps{0}, is rejected together with all posts in sequence, including those in
+the common suffix.
+%
+This rejection rule for remaining posts is a direct consequence of tamper proof
+Merkle~DAGs~\cite{p2p.ipfs,p2p.bitcoin} typically used in permissionless
+protocols, which cannot support graph modifications.
 
@@
 %
-Peers that first received branches with less reputation will need to reorder
-all posts starting at the forking point.
+Peers that first received branches with less reputation need to reorder all
+posts starting at the forking point.
 This might involve removing content in the end-user software.
@@
     (iii) is based on reputation of an agreed common prefix,
-    (iv) leading to a global and deterministic and result.
-%
-Note that the consensus order exists only for accountability purposes, and is
-a view of the DAG structure, which remains as the primary source of data.
+    (iv) leading to a global and deterministic result.
+%
+Note that the consensus order exists for reputation accountability purposes,
+and is only a view of the primary DAG structure.
 
 \subsection{Malicious Behaviors and Hard Forks}
+\label{sec.design.hard}
 
 Now, consider a malicious scenario in which user $c$ in Figure~\ref{fig.reps}
-intentionally disconnected from the network for weeks to cultivate fake
+intentionally disconnects from the network for weeks to cultivate fake
 identities $x$ and $y$ and accumulate \reps.
-When merging, these fake users become legit and now constitute the majority of
-the forum in terms of reputation.
-%
-However, because the common prefix is the basis of consensus, users in the
-branch with more reputation can still react even after the attack.
+When merging, these fake users become legitimate and now accumulate the
+majority of \reps in the forum.
+%
+However, this kind of attack is refutable because, since the common prefix is
+the basis of consensus, users in the branch with more reputation can still
+react even a posteriori.
 For instance, users $a$ and $b$ can pretend that they did not yet see
-\code{branch-1} and post extra dislikes to user $c$ from \code{branch-2} before
-the common suffix so that a further merge removes all posts of \code{branch-1}
-from the DAG.
-%
-The example shows that this kind of attack is innocuous because it requires at
-least half of the \reps accumulated in the common prefix.
-%
-The same countermeasure can be used against a minority of users colluding to
-boost or protect their own posts.
+\code{branch-1} and post extra dislikes to user $c$ in \code{branch-2} so that
+a further merge invalidates all posts from \code{branch-1}, removing them from
+the DAG.
+%
+This kind of attack is innocuous because it requires at least half of the \reps
+accumulated in the common prefix.
+%
+The same countermeasure can be used against a minority of malicious users
+colluding to boost or protect their own posts.
 
@@
 As a counterpoint, maybe users $a$ and $b$ have abandoned the forum for months,
-and thus \code{branch-1} is actually legit.
+and thus \code{branch-1} is actually legitimate.
 In this case, users $a$ and $b$ might be the ones trying to take over the
-forum.
-Yet another possibility is that both branches are legit but became disconnected
-for a long period of time.
-In any case, it is unacceptable for users that a very old remote branch affects
-a long active forum.
-%
+forum and would succeed, since they control the majority of \reps in the common
+prefix.
+Yet another possibility is that both branches are legitimate but became
+disconnected for a long period of time.
+In any case, it is unacceptable for correct users that a very old remote branch
+affects a long active forum.
+
 For this reason, the consensus algorithm includes an extra constraint that
@@
 takes priority and is ordered first.
-This situation is analogous to a hard fork in Bitcoin and the branches will
-never synchronize again.
+This situation is analogous to a hard fork in Bitcoin and the branches become
+incompatible and will never synchronize again.
 More than simple numeric disputes, a hard fork represents a social conflict in
 which reconciling branches is no longer possible.
-%
+
 Figure~\ref{fig.hard} illustrates hard forks by distinguishing \emph{stable
@@
     \item \code{i} or \code{j} is first, whichever has more reputation in the
-          common prefix.
+          common prefix, as discussed in Section~\ref{sec.design.basic}.
     \item otherwise, branches are ordered by an arbitrary criteria, such as
@@
 blockchain must prevail, public forums can split and diverge indiscriminately.
-For instance, users can explicitly apply hard forks at any point in time or
+For instance, users can explicitly apply hard forks at any point in time to
 ``reboot'' a community from a previous state.
 Ultimately, the consensus algorithm provides a transparent mechanism to help
-users understand the evolution of public forums and act accordingly.
+users understand the evolution of public forums and act accordingly, even if it
+leads to hard forks.
 
@@
 social content publishing, and thus, an important contribution of this work.
+
 As described in Figure~\ref{fig.state}, a forum post has three possible states:
@@
 Otherwise, it is \code{BLOCKED} and requires a like from another user.
+
 Blocked posts are not considered part of the forum DAG in the sense that new
@@
 chance to be welcomed with a like.
-A reasonable policy is to hold blocked posts in a temporary bag and retransmit
-them for some visibility in the network.
-%
+A reasonable policy is to hold blocked posts in a temporary but public bag and
+retransmit them for some visibility in the network.
+
 Once accepted, a post becomes part of the forum and can never be removed
@@
 However, if the number of dislikes exceeds an arbitrary threshold (e.g., more
-dislike than likes), the post becomes \code{REVOKED} and its payload is not
+dislikes than likes), the post becomes \code{REVOKED} and its payload is not
 retransmitted to other peers.
@@
 Later, if the post receives new likes, it means that the payload is still known
-somewhere and peers can request it when synchronizing again.
-
-\subsection{Synchronization and Byzantine Peers}
+somewhere and peers can request it when synchronizing again, since the post
+becomes \code{ACCEPTED} again.
+
+\subsection{Peer Synchronization and Byzantine Faults}
 
@@
 the messages.
-When peers synchronize, they mutually synchronize missing branches such that
-they reach the same DAG structure~\cite{p2p.byz}.
+When peers synchronize, they mutually exchange missing branches such that they
+reach the same DAG structure~\cite{p2p.byz}.
 As they synchronize, each peer runs the consensus algorithm locally and
@@
 Because the protocol relies on a tamper-proof DAGs, and because each peer
-self-verifies the posts as they synchronize, the protocol becomes tolerant to
-Byzantine-faulty peers~\cite{lamport.byz}.
-%
-No information coming from any peer is trusted a priori.
+self verifies the posts as they synchronize, the protocol becomes tolerant to
+Byzantine faults~\cite{lamport.byz}.
+%
+Note that no information coming from any peer is trusted a priori.
 %
@@
 
-Therefore, although Byzantine peers may still perform some kind of denial of
+Therefore, although Byzantine peers may still perform some sort of denial of
 service attacks in the network, they cannot modify the DAG structure, create
-new content (without performing work), or impede that two correct nodes
+arbitrary content (without performing work), or impede that two correct nodes
 synchronize directly.
@@
 support content moderation and mitigate abuse.
-Table~\ref{fig.rules} details the rules we conceived, which are discussed as
-follows through a concrete example.
+Table~\ref{fig.rules} details the concrete reputation rules we conceived for
+\FC, which is compatible with the general rules of Table~\ref{fig.general}, and
+are discussed as follows through an example.
 Authors have to sign posts in order to be accounted by the reputation system
@@
 
-{\scriptsize
+{\footnotesize
 \begin{verbatim}
@@
  10AE3E..           <-- hash representing the chain
- > freechains 'forum' post --sign='DA3B5F..' \
-    'The purpose of this chain is...'
+ > freechains 'forum' post --sign='DA3B5F..' 'The chain purpose is...'
  1_CC2184..         <-- hash representing the post
@@
 assigning \nreps{30} equally distributed between an arbitrary number of
-pioneers indicated through their public keys.
+pioneers indicated through their public keys (one in this example).
 The pioneers shape the initial culture of the chain with the first posts and
@@
 %
-The \code{post} command in sequence is signed by the single pioneer (in this
-example) and indicates the purpose of the chain for future users.
+The \code{post} command in sequence is signed by the single pioneer and
+indicates the purpose of the chain for future users.
 
@@
 chains.
-Fully peer-to-peer systems cannot rely on logins or {\scriptsize CAPTCHAs}
+Fully peer-to-peer systems cannot rely on logins or {\footnotesize CAPTCHAs}
 due to the lack of a central authority.
@@
  10AE3E..           <-- same pioneer as before
- > freechains 'forum' post 'Im a newbie...' \
-    --sign='41DDF1..'
+ > freechains 'forum' post 'Im a newbie...' --sign='41DDF1..'
  2_C3A40F..         <-- blocked post
- > freechains 'forum' like '2_C3A40F..' \
-    --sign='DA3B5F..'
+ > freechains 'forum' like '2_C3A40F..' --sign='DA3B5F..'
  3_59F3E1..         <-- hash representing the like
@@
 Note that chains with the same name but different pioneers would be
-incompatible because the hash of genesis posts depend on the pioneers' public
+incompatible because the hash of genesis posts depends on the pioneers' public
 keys.
@@
 Back to Figure~\ref{fig.forum}, a new user with \nreps{0} tries to post a
-message (hash~\code{C3A40F..}) and is blocked (rule~\code{4.a}), as the red
+message (hash~\code{2\_C3A40F..}) and is blocked (rule~\code{4.a}), as the red
 background highlights.
-But the pioneer liked the blocked message, decreasing herself to \nreps{29}
+But the pioneer likes the blocked message, decreasing herself to \nreps{29}
 and increasing new user to \onerep (rule~\code{3.a}).
@@
 \nreps{30} and the new user \nreps{2}, growing the economy in \nreps{2} as
-result of the two consolidated posts.
+result of the two new consolidated posts.
 Note that rule~\code{1.b} rewards at most one post of each author at a time.
-Hence, new posts during the 24-hour period will not reward each of them with
-extra \reps.
+Hence, extra posts during the 24-hour period do not reward authors with extra
+\reps.
 Note also that rule~\code{4.b} limits authors to at most \nreps{30}, which
@@
 and identities behind posts are not directly assessed by the protocol, but
-instead by the other users in the system.
+instead by the judgement of other users in the system.
 %
@@
 
-\section{Experiments with Public Forums}
+\subsection{Experiments with Public Forums in \FC}
 \label{sec.evaluation}
@@
 In this section, we perform experiments to evaluate the performance and
-practicability of the protocol.
+practicability of \FC and its consensus algorithm.
 %
@@
 For each message in the original forum, we
-    (i)   set the timestamp of all peers to match the date,
+    (i)   set the timestamp of all peers to match the original date,
     (ii)  create a pair of keys if the author is new,
@@
 Since all messages are part of the original archive, we always perform step
-(iv) to unblock messages from newbies, which we use to measure the evaluation
-parameter (d).
+(iv) to unblock messages from newbies, which we also use to measure the
+evaluation parameter (d).
 Step (v) will inevitably create forks in the chain for any \texttt{M<N}, which
@@
 %
-The incremental sort is limited to 7 days or 100 posts, regardless of the size
-of the chain, which conveniently settles an upper bound on the input size of
-the consensus algorithm.
+The incremental sort is limited to 7 days or 100 posts by design (as discussed
+in Section~\ref{sec.design.hard}), regardless of the size of the chain, which
+conveniently settles an upper bound on the input size of the consensus
+algorithm.
 %
@@
 Evaluation item (c) counts the number of forks in chain DAGs, which indicates
-the level of asynchrony between peers following the local-first principle
-(supposedly).
+the level of asynchrony between peers following the local-first principle.
 We calculate the ratio of forks over the total number of messages, e.g., a DAG
@@
 Even though the newbie rule~\code{4.a} in Table~\ref{fig.rules} is key to
-combat Sybils, ideally it should not deny access to active users with low
-reputation recurrently.
+combat Sybils, ideally it should not recurrently deny access to active users
+with low reputation.
 %
@@
 As an example, if $10$ users posted $110$ messages requiring $20$ likes, we
-discount the $10$ initial messages and welcoming likes and find a ratio of
-$10\%$ (\code{(20-10)/(110-10)}).
+discount the $10$ initial messages and welcoming likes (when the user first
+appears), and find a ratio of $10\%$ (\code{(20-10)/(110-10)}).
 %
@@
 architectures compared to peer-to-peer systems, which typically include message
-signing, hash linking, and extra verification rules.
+signing, hash linking, and extra verification rules (as evaluated in
+Section~\ref{sec.evaluation}).
 
@@
 Scuttlebutt~\cite{p2p.scuttlebutt} and Aether~\cite{p2p.ecosystem} are closer
-to \FC goals and cover human-centric \Xon and \Xnn public communication,
-respectively~\cite{fcs.sbseg20}.
+to \FC goals and cover human communication.
 
@@
 locally.
-Scuttlebutt is aligned to \Xon public identity chains of \FC.
-%
-For group \Xnn communication, Scuttlebutt uses the concept of channels, which
-are in fact nothing more than hash tags (e.g. \emph{\#sports}).
+Note however that the basic unit of communication in Scuttlebutt is
+unidirectional , with a public identity broadcasting a feed to its
+followers (\Xon).
+%
+For open groups communication (\Xnn), Scuttlebutt uses the concept of channels,
+which are in fact nothing more than hash tags (e.g. \emph{\#sports}).
 Authors can tag posts, which appear not only in their feeds but also in local
@@
 Aether~\cite{p2p.ecosystem} provides peer-to-peer public communities aligned
-with \Xnn public forums of \FC.
+with public forums of \FC (\Xnn).
 A fundamental difference is that Aether is designed for ephemeral, mutable
@@
 for social content sharing in peer-to-peer networks.
-We enumerate four main contributions:
+We enumerate three main contributions:
     (i)   human authored content as a scarce resource
@@
 The reputation and consensus mechanism is integrated into \FC, a peer-to-peer
-protocol that offers multiple arrangements of public and private
-communications.
-Users have the power to create public forums of interest and apply diverse
+protocol, in which users can create public forums of interest and apply diverse
 moderation policies.
@@
 proof-of-work), which do not appraise social interactions and also tend
-to concentrate over the time.
+to concentrate power over the time.
 %

### 2023-06-08 freechains/freechains.tex
...

@@
 increasingly more centralized in a few companies (e.g. Facebook and
-Twitter)~\cite{internet.fixing,p2p.osn,p2p.dosn}.
+Twitter)~\cite{internet.fixing}.
+%,p2p.osn,p2p.dosn}.
 %These companies benefit from closed protocols and network effects to keep
@@
 (iii) are respectful and on topic.
-In a centralized system, items (i) and (ii) are trivially achieved assuming
-availability and delivery order in the service, while for item (iii), users
-have to trust the service to moderate content.
-In a decentralized setting, however, none of these demands are easily
+In centralized systems, items (i) and (ii) are trivially achieved assuming
+availability and delivery order, while item (iii) assumes that users trust the
+service to moderate content.
+In decentralized systems, however, none of these demands are easily
 accomplished.
-A common approach in gossiping protocols is to proactively replicate and
-disseminate posts among peers until they reach all
-users~\cite{p2p.survey,p2p.byz}.
+A common approach in gossiping protocols is to proactively disseminate posts
+among peers until they reach all users~\cite{p2p.survey,p2p.byz}.
 However, this approach does not guarantee consensus since posts can be received
-in conflicting orders~\cite{p2p.intention,p2p.dvcs}.
+in conflicting orders~\cite{p2p.intention}.
+%,p2p.dvcs}.
 %As an example, antagonistic messages such as \emph{"X is final"} vs
@@
 For instance, Bitcoin transactions permanently store arbitrary human-authored
-content~\cite{btc.data}, including harmful texts and images~\cite{btc.content}.
+content, including harmful texts and images~\cite{btc.content}.
+%~\cite{btc.data}
 
@@
 We integrated the proposed consensus algorithm into \FC~\cite{fcs.sbseg20}, a
-practical peer-to-peer content dissemination protocol that provides strong
-eventual consistency~\cite{p2p.crdts,p2p.sec}.
-%
+practical peer-to-peer (P2P) content dissemination protocol that provides
+strong eventual consistency~\cite{p2p.sec}.
+%p2p.crdts,
 To show the practicability of the consensus mechanism, we simulated months of
@@
 In Section~\ref{sec.related}, we compare our system with publish-subscribe
-protocols, federated applications, and fully peer-to-peer systems.
+protocols, federated applications, and fully P2P systems.
 In Section~\ref{sec.conclusion}, we conclude this work.
@@
 
-In the absence of moderation, permissionless peer-to-peer forums are
-impractical, mostly because of Sybils abusing the system.
+In the absence of moderation, permissionless P2P forums are impractical, mostly
+because of Sybils abusing the system.
 For instance, it should take a few seconds to generate thousands of fake
-identities and SPAM millions of messages into the system.
+identities and SPAM millions of messages.
 %Hence, without moderation, there are no limits on the number and size of posts
@@
 \centering
-\includegraphics[width=0.75\textwidth]{general.png}
+\includegraphics[width=0.65\textwidth]{general.png}
 \caption{General reputation rules in public forums.}
@@
 Our key idea to stablish consensus in forum DAGs is to favor forks with posts
-from users that constitute the majority of the reputation in the network.
-These forks have more associated work from active users and are analogous to
-longest chains in Bitcoin.
+from users that constitute the majority of the reputation.
+These forks have more associated work and are analogous to longest chains in
+Bitcoin.
 %
@@
 %
-We assume that within the common prefix, users $a$ and $b$ have contributed
-with better content and therefore have more reputation combined than $c$ has
-alone (i.e., $8+5>3$).
-We will use this information when ordering branches in the consensus algorithm.
-%
-Then, user $c$ disconnects from $a$ and $b$, and evolves in separate along with
-new users $x$ and $y$ with posts in \code{branch-1}.
+We assume that within the prefix, users $a$ and $b$ have contributed with
+better content and therefore have more reputation combined than $c$ has alone
+(i.e., $8+5>3$).
+We will use this fact when ordering branches in the consensus algorithm.
+%
+Then, user $c$ disconnects from $a$ and $b$, and evolves with new users $x$ and
+$y$ with posts in \code{branch-1}.
 In the meantime, $a$ and $b$ participate with posts in \code{branch-2}.
@@
 possibly block posts from users in the other branch, depending on how they are
-ordered.
-Therefore, we need to order the branches with a global consensus, such that we
-can validate operations consistently across the network.
+finally ordered.
+Therefore, we need to order branches with a global consensus, such that we can
+validate operations consistently across the network.
 %
@@
 Therefore, based on the reputation of users in the common prefix,
-Figure~\ref{fig.reps}.B indicates consensus order between posts in the forum
+Figure~\ref{fig.reps}.B indicates the consensus order between posts in the forum
 DAG:
     the common prefix comes first,
-    followed by all posts in \code{branch-2},
-    followed by all posts in \code{branch-1},
-    followed by the common suffix.
+    then all posts in \code{branch-2},
+    then all posts in \code{branch-1},
+    then the common suffix.
 %
@@
 %
-Peers that first received branches with less reputation need to reorder all
-posts starting at the forking point.
+Peers that first receive branches with less reputation need to reorder all
+posts after the forking point.
 This might involve removing content in the end-user software.
-This behavior is similar to Bitcoin's blockchain reorganization, when a peer
-detects a new longest chain. % and disconsiders old blocks.
-%
-Likewise, peers that first saw branches with more reputation just need to put
-the other branch in sequence with no reordering at all.
+This behavior is similar to Bitcoin's blockchain reorganization, when peers
+detect new longest chains. % and disconsiders old blocks.
+%
+Likewise, peers that first saw branches with more reputation just need to
+append the other branch, with no reordering at all.
 %and do not need to recompute anything.
@@
     (ii) require consensus to order operations, which
-    (iii) is based on reputation of an agreed common prefix,
+    (iii) is based on a common prefix reputation,
     (iv) leading to a global and deterministic result.
 %
-Note that the consensus order exists for reputation accountability purposes,
-and is only a view of the primary DAG structure.
+Note that the consensus order exists only to account reputation and verify
+operations, and is only a view of the primary DAG structure.
 
@@
 
-Now, consider a malicious scenario in which user $c$ in Figure~\ref{fig.reps}
+Consider a malicious scenario in which user $c$ in Figure~\ref{fig.reps}
 intentionally disconnects from the network for weeks to cultivate fake
-identities $x$ and $y$ and accumulate \reps.
-When merging, these fake users become legitimate and now accumulate the
+identities $x$ and $y$ with the intention to accumulate \reps.
+When merging, these fake users would become legitimate and accumulate the
 majority of \reps in the forum.
@@
 For instance, users $a$ and $b$ can pretend that they did not yet see
-\code{branch-1} and post extra dislikes to user $c$ in \code{branch-2} so that
-a further merge invalidates all posts from \code{branch-1}, removing them from
-the DAG.
+\code{branch-1} and post extra dislikes to user $c$ so that a further merge
+invalidates all posts from \code{branch-1}, removing them from the DAG.
 %
@@
 \centering
-\includegraphics[width=\textwidth]{n100-d7.png}
+\includegraphics[width=0.85\textwidth]{n100-d7.png}
 \caption{
-    Stable consensus permanently freezes the order of posts once they cross
-    the threshold.
+    Stable consensus freezes the order of posts once they cross the threshold.
     The unstable order may still be affected by incoming branches.
@@
 \centering
-\includegraphics[width=0.75\textwidth]{state-revoked.png}
+\includegraphics[width=0.65\textwidth]{state-revoked.png}
 \caption{
@@
 posts do not link back to it.
-In addition, peers are not required to hold blocked posts and neither
-retransmit them to other peers.
+In addition, peers are not required to hold blocked posts nor retransmit them
+to other peers.
 However, if blocked posts are not disseminated, new users will never have the
 chance to be welcomed with a like.
-A reasonable policy is to hold blocked posts in a temporary but public bag and
-retransmit them for some visibility in the network.
+Therefore, a reasonable policy is to hold blocked posts in a temporary but
+public bag and retransmit them for some visibility.
 
@@
 
-\FC~\cite{fcs.sbseg20} is an unstructured peer-to-peer topic-based
-publish-subscribe protocol, in which each \emph{chain} is a replicated
-Merkle-DAG representing the causal relationships between the messages.
+\FC~\cite{fcs.sbseg20} is an unstructured P2P topic-based publish-subscribe
+protocol, in which each \emph{chain} is a replicated Merkle-DAG representing
+the causal relationships between the messages.
 %
@@
 
-The most basic concern in public forums is to resist Sybils abusing the
-chains.
-Fully peer-to-peer systems cannot rely on logins or {\footnotesize CAPTCHAs}
-due to the lack of a central authority.
+The most basic concern in public forums is to resist Sybil attacks.
+Fully P2P systems cannot rely on logins or {\footnotesize CAPTCHAs} due to the
+lack of a central authority.
 Viable alternatives include (i) building social trust graphs, in which users
@@
 \centering
-\includegraphics[width=0.75\textwidth]{forum.png}
+\includegraphics[width=0.65\textwidth]{forum.png}
 \caption{
@@
 
-Likes and dislikes (rules \code{3.a} and \code{3.b}) serve three purposes
-in the chains:
+Likes and dislikes (rules \code{3.a} and \code{3.b}) serve three purposes:
     (i) welcoming new users,
     (ii) measuring the quality of posts, and
-    (iii) revoking abusive posts (SPAM, fake news, illicit content, etc).
+    (iii) revoking abusive posts (SPAM, fake news, etc).
 %
@@
 
-In this section, we perform experiments to evaluate the performance and
-practicability of \FC and its consensus algorithm.
+We performed experiments to evaluate the performance of \FC and its consensus
+algorithm.
 %
@@
 nature of the protocol.
-For the chat, we use \texttt{N=5} and \texttt{M=3}, which represents a smaller
-number of peers with more interconnections.
-%
-We executed each simulation $4$ times in a conventional desktop PC (\texttt{i7}
-CPU, \texttt{8GB} RAM, \texttt{512GB} SSD).
+For the chat, we use \texttt{N=5} and \texttt{M=3}, which is a smaller number
+of peers but with more interconnections.
+%
+We executed each simulation $4$ times in a desktop PC (\texttt{i7} CPU,
+\texttt{8GB} RAM, \texttt{512GB} SSD).
 Since the variations were negligible, we always discuss the median measures.
@@
 sharing~\cite{p2p.survey,p2p.ecosystem}.
-Here we consider the classes of publish-subscribe, federated, and peer-to-peer
+Here we consider the classes of publish-subscribe, federated, and P2P
 protocols.
@@
 Nevertheless, some pubsubs do not rely on server roles, and instead, use
-peer-to-peer gossip dissemination~\cite{pubsub.tera,pubsub.rappel,pubsub.stan,pubsub.vitis,pubsub.gossipsub,pubsub.rappel}.
+P2P gossip dissemination~\cite{pubsub.tera,pubsub.rappel,pubsub.gossipsub}.
+%,pubsub.stan,pubsub.vitis
 Most of these protocols focus on techniques to achieve scalability and
@@
 
-Federated protocols, such as e-mail, allow users from one domain to seamlessly
-exchange messages with users of other domains.
+Federated protocols, such as e-mail, allow users from one domain to exchange
+messages with users of other domains.
 \emph{Diaspora}, \emph{Matrix}, and \emph{Mastodon} are recent federations for
@@
 The number of hops and header overhead can be much smaller in client-server
-architectures compared to peer-to-peer systems, which typically include message
-signing, hash linking, and extra verification rules (as evaluated in
+architectures compared to P2P systems, which typically include message signing,
+hash linking, and extra verification rules (as evaluated in
 Section~\ref{sec.evaluation}).
@@
 
-Regarding peer-to-peer protocols, Bitcoin~\cite{p2p.bitcoin} is probably the
-most successful permissionless network, but serves specifically for electronic
+Regarding P2P protocols, Bitcoin~\cite{p2p.bitcoin} is probably the most
+successful permissionless network, but serves specifically for electronic
 cash.
@@
 As an advantage, these proof mechanisms are generic and apply to multiple
-domains, since they depend on an extrinsic scarce resource.
+domains, since they depend on extrinsic resources.
 % (i.e., the richer gets richer)
@@
 Dat~\cite{p2p.dat} around mutable pubkey-addressed data.
-IPFS is more suitable to share large and stable content such as movies and
-archives, while Dat is more suitable for dynamic content such as web apps.
+IPFS is more suitable to share large and stable content such as movies, while
+Dat is more suitable for dynamic content such as web apps.
 %
@@
 
-Aether~\cite{p2p.ecosystem} provides peer-to-peer public communities aligned
-with public forums of \FC (\Xnn).
+Aether~\cite{p2p.ecosystem} provides P2P public communities aligned with public
+forums of \FC (\Xnn).
 A fundamental difference is that Aether is designed for ephemeral, mutable
@@
 In this paper, we propose a permissionless consensus and reputation mechanism
-for social content sharing in peer-to-peer networks.
+for social content sharing in P2P networks.
 We enumerate three main contributions:
@@
 
-The reputation and consensus mechanism is integrated into \FC, a peer-to-peer
-protocol, in which users can create public forums of interest and apply diverse
+The reputation and consensus mechanism is integrated into \FC, a P2P protocol,
+in which users can create public forums of interest and apply diverse
 moderation policies.

### 2023-06-08 freechains/freechains.tex
...

@@
 and robust access.
-On the other hand, they concentrate too much power over its users, holding and
-monetizing over their private data and behavior.
+On the other hand, they concentrate excessive power, controlling and monetizing
+over private user data.
 %
@@
 (i)   reach all users,
-(ii)  are delivered in a consistent order, and
+(ii)  are delivered in consistent order, and
 (iii) are respectful and on topic.
@@
 content in the middle of a blockchain, which is inadmissible considering
-illicit content (e.g., hate speech).
-For instance, Bitcoin transactions permanently store arbitrary human-authored
-content, including harmful texts and images~\cite{btc.content}.
+illicit content (e.g., hate speech)~\cite{btc.content}.
+%For instance, Bitcoin transactions permanently store arbitrary human-authored
+%content, including harmful texts and images.
 %~\cite{btc.data}
@@
 With such \emph{proof-of-authoring} mechanism, token generation is expensive,
-while verification is cheap and made by multiple users.
+while verification is cheap and shared by multiple users.
 %
@@
 
-\begin{table}
-\centering
-\includegraphics[width=0.65\textwidth]{general.png}
-\caption{General reputation rules in public forums.}
-\label{fig.general}
-\end{table}
-
-We propose a reputation system in which users can spend tokens named \reps to
-post and rate content in the forums:
+In our system, users can spend tokens named \reps to post and rate content in
+forums:
 a \code{post} initially penalizes authors until it consolidates and counts
@@
 \footnote{
-    Note that to some extent AI content may be indistinguishable from human
-    content, but its frequency and relevance in forums is still subject to
-    human evaluation in our proposal.
+    Note that although human and AI content may become indistinguishable, their
+    frequency and relevance in forums is still subject to human evaluation in
+    our proposal.
 }
+
+\begin{table}
+\centering
+\includegraphics[width=0.65\textwidth]{general.png}
+\caption{General reputation rules in public forums.}
+\label{fig.general}
+\end{table}
 
@@
 Therefore, in order to impose scarcity, we determine that only content
-authoring generates \reps, while likes and dislikes just transfer \reps between
+authoring generates \reps, while likes and dislikes only transfer \reps between
 users.
@@
 
+Our key idea to stablish consensus in forum DAGs is to favor forks with posts
+from users that constitute the majority of the reputation.
+These forks have more associated work and are analogous to longest chains in
+Bitcoin.
+%
+In technical terms, we adapt a topological sorting algorithm to favor
+reputation when deciding between branches in a forum DAG.
+
+Figure~\ref{fig.reps}.A illustrates the evolution of a forum DAG in the
+presence of a network partition:
+A common prefix has signed posts from users $a$, $b$, and $c$, when they were
+still all connected.
+%
+We assume that within the prefix, users $a$ and $b$ have contributed with
+better content and therefore have more reputation combined than $c$ has alone
+(i.e., $8+5>3$).
+We will use this fact to order branches in the consensus algorithm.
+%
+Then, user $c$ disconnects from $a$ and $b$, and evolves with new users $x$ and
+$y$ in \code{branch-1}.
+In the meantime, $a$ and $b$ participate with new posts in \code{branch-2}.
+After some time, the partitions reconnect and new posts from $x$, $b$ and $c$
+merge the branches now with a common suffix.
+
 \begin{figure}
@@
 
-Our key idea to stablish consensus in forum DAGs is to favor forks with posts
-from users that constitute the majority of the reputation.
-These forks have more associated work and are analogous to longest chains in
-Bitcoin.
-%
-In technical terms, we adapt a topological sorting algorithm to favor
-reputation when deciding between branches in a forum DAG.
-
-Figure~\ref{fig.reps}.A illustrates the evolution of a forum DAG in the
-presence of a network partition:
-A common prefix has signed posts from users $a$, $b$, and $c$, when they were
-still all connected.
-%
-We assume that within the prefix, users $a$ and $b$ have contributed with
-better content and therefore have more reputation combined than $c$ has alone
-(i.e., $8+5>3$).
-We will use this fact when ordering branches in the consensus algorithm.
-%
-Then, user $c$ disconnects from $a$ and $b$, and evolves with new users $x$ and
-$y$ with posts in \code{branch-1}.
-In the meantime, $a$ and $b$ participate with posts in \code{branch-2}.
-After some time, the partitions reconnect and new posts from $x$, $b$ and $c$
-merge the branches with a common suffix.
-
 Note that even non-malicious network partitions may experience conflicts when
 merging branches.
-For instance, dislike operations in one branch to common past posts could
-possibly block posts from users in the other branch, depending on how they are
-finally ordered.
+For instance, dislike operations in one branch to past posts in the common
+prefix could possibly block posts from users in the other branch, depending on
+how they posts are ordered.
 Therefore, we need to order branches with a global consensus, such that we can
@@
 
-Therefore, based on the reputation of users in the common prefix,
-Figure~\ref{fig.reps}.B indicates the consensus order between posts in the forum
-DAG:
+Based on the reputation of users in the common prefix, Figure~\ref{fig.reps}.B
+indicates the consensus order between posts in the forum DAG:
     the common prefix comes first,
@@
 However, the longer a peer remains disconnected, the more conflicting
-operations it may see, and the higher are the chances of reordering when
+operations it may see, and the higher are the chances of posts reordering when
 rejoining.
@@
 Note that the consensus order exists only to account reputation and verify
-operations, and is only a view of the primary DAG structure.
+operations, and is only a view of the primary DAG structure of public forums.
 
@@
 disconnected for a long period of time.
-In any case, it is unacceptable for correct users that a very old remote branch
-affects a long active forum.
+In any case, it is unacceptable that a very old remote branch affects a long
+active forum.
 
@@
 
-In summary, the rules to merge a branch from a remote machine \code{j} into a
-branch from a local machine \code{i} are as follows:
+In summary, the rules to merge a branch \code{j} from a remote machine into a
+branch \code{i} from a local machine are as follows:
 \begin{itemize}
@@
           common prefix, as discussed in Section~\ref{sec.design.basic}.
-    \item otherwise, branches are ordered by an arbitrary criteria, such as
-          lexicographical order of the post hashes immediately after the
-          common prefix.
+    \item otherwise, branches use the arbitrary criteria of lexicographical
+          order of the post hashes immediately after the common prefix.
 \end{itemize}
@@
 
-Because the protocol relies on a tamper-proof DAGs, and because each peer
-self verifies the posts as they synchronize, the protocol becomes tolerant to
+Because the protocol relies on tamper-proof DAGs, and because each peer self
+verifies the posts as they synchronize, the protocol becomes tolerant to
 Byzantine faults~\cite{lamport.byz}.
@@
 Therefore, although Byzantine peers may still perform some sort of denial of
-service attacks in the network, they cannot modify the DAG structure, create
-arbitrary content (without performing work), or impede that two correct nodes
-synchronize directly.
+service attacks in the network, they cannot modify the DAG structure, nor
+create arbitrary content, nor impede that two correct nodes synchronize
+directly.
 
@@
 protocol, in which each \emph{chain} is a replicated Merkle-DAG representing
-the causal relationships between the messages.
+the causal relationships between posts.
 %
 The protocol operation is typical of publish-subscribe systems: an author
-publishes a post to a chain, and subscribed users eventually receive the
-message.
+publishes a post to a chain, which subscribed users eventually receive.
 % by synchronizing DAGs.
@@
 
-We integrated the proposed reputation and consensus mechanism with \FC to
-support content moderation and mitigate abuse.
+In order to support content moderation and mitigate abuse, we integrated the
+proposed reputation and consensus mechanism of Section~\ref{sec.design} with
+\FC.
 Table~\ref{fig.rules} details the concrete reputation rules we conceived for
 \FC, which is compatible with the general rules of Table~\ref{fig.general}, and
-are discussed as follows through an example.
+which are discussed as follows through an example.
 Authors have to sign posts in order to be accounted by the reputation system

### 2023-08-09 freechains/rebuttal.md
...

# Rebuttal Letter

We would like to thank the reviewers for the thoughtful comments, questions,
and criticism.
We will accommodate all clarifications and answers in a final version of the
paper.
Each comment is addressed individually as follows.

## Reviewer 1

> Quanto a solução, senti falta de alguma discussão sobre as limitações.
> Em especial, a falta de finalidade, i.e., um post pode ser excluído devido
> aos forks (no momento do merge).
> Este post poderia ser de um usuário legitimo ? quanto isso ocorre e quais as
> implicações?

Sim, para isso, ele teria que escrever em um fork "obscuro", desconectado do
resto dos usuários legítimos.
Ao ser reconectado, esse fork poderia sim ter posts removidos.
A única implicação é que ele teria que postar novamente a mesma mensagem.

No meio da Seção 2.1, essa possibilidade é levantada, mas não com tanta ênfase:

> While applying the branches in order, if any post operation fails, all
> remaining posts are rejected and removed from the DAG, as if they never
> existed.

Em termos práticos, os usuários legítimos se conectam entre si ao longo do
tempo, o que deveria ser suficiente para evitar essa situação.
Para a versão final, nós adicionamos o seguinte trecho:

> Legitimate posts in secondary branches might be removed due to merges, which
> requires that users repost the messages. Nevertheless, we expect that
> legitimate users gossip among themselves more often, preventing this
> situation.

> Além disso, é possível que um usuário malicioso participe de mais de uma
> cadeia em um fork? neste caso como fica a regra de merge? e se der empate,
> todos devem escolher o mesmo fork para começar, como isso seria feito?

A reputação de um usuário em uma cadeia X não é válida em uma cadeia Y.
Portanto, interpretamos o comentário acima como "...participe de mais de um
fork em uma mesma cadeia".

Sim, se o usuário malicioso dispor de 1 rep, ele pode escrever propositalmente
em vários forks.
Nesse caso, a reputação prévia dele vai contar igualmente em todos os merges,
se anulando e, portanto, não influenciando no resultado final do merge.

Em caso de empate, que pode acontecer mesmo sem usuários maliciosos, o hash do
fork será critério considerado por todos igualmente.
Esse critério é descrito no final da Seção 2.2:

> ...otherwise, branches use the arbitrary criteria of lexicographical order of
> the post hashes immediately after the common prefix.

Agora vamos discutir um possível ataque usando essa estratégia.
O usuário malicioso poderia escrever em vários branches antes de usuários
legítimos, de forma que os merges acabariam por remover mensagens legítimas.
Essa situação é análoga ao comentário anterior e obrigaria os usuários a
repostarem as mensagens.
Nós entendemos que os usuários legítimos estarão mais conectados,
inviabilizando esse tipo de ataque.

> Pela descrição, aparentemente na Figura 3 deveria ter uma transição de
> revoked  para accepted.

A Figura 3 já tem essa transição.
Há uma seta bi-direcional entre os dois estados.
(Espero ter entendido corretamente esse comentário.)

# Reviewer 2

> It is strange that known techniques like EigenTrust are not even mentioned,
> even though we are in the blockchain domain.

> We do not know if the results are good or bad because there was no comparison
> to the state of the art or baseline.

The main goal of our experiments in Section 3.1 is to stress the consensus
algorithm to show that the mechanism is practical, i.e., that it has a
reasonable performance.
The good result is that the protocol works in a permissionless context,
regardless of the slower performance.

We added the following remark at the beginning of Section 3.1:

> The main goal of this section is to stress the consensus algorithm to show
> that permissionless public forums are viable, regardless of a slower
> performance in comparison to permissioned protocols.

> Also, at the end of the day, we do not know if the proposal is resistant to
> Sybil attacks because there was no validation of the consensus and reputation
> mechanisms.

The protocol is Sybil resistant because it is expensive to write to a chain.
By default, users with no reps do not have write access, as described in the
beginning of Section 2:

> To prevent Sybils, users with no *reps* cannot operate under these rules,
> requiring a welcoming like from any other user already in the system.
> The fact that likes are zero-sum operations, which only transfer reputation,
> eliminates the incentives from malicious users to invite Sybils into the
> system.
> The only way to generate new *reps* is to post content that other users
> tolerate, which demands non-trivial work resistant to automation.

We included the following remarks:

> In this sense, having one or hunderds of machines in the network does not
> affect the capacity to write into the chains.
> ...
> In our analogy with Bitcoin, Sybil attacks would require to add human
> resurces, instead of CPU power.

# Reviewer 3

> It's important to note that the reputation system does not enforce "good"
> human behavior but provides a transparent and quantitative way for users to
> understand forum evolution and act accordingly.

> I am also concerned that the content removal is too weak to work in practice.
> For example, a Forum may be taken over by extremists, and they can upvote
> illegal content. That would require removal due to legal process, or face
> blocking at network level.

This is all true, and the reason why we close the paper with your remark above.
It is not even required to be "taken by extremists", but only "created by and
for extremists".
But we also believe that a protocol that is tamper proof and provides
persistence can actually help law enforcement.

> Maybe a way to go is to create a pool of super-reputations that could answer
> to injunctions and administratively remove the content and a measure to avoid
> illegal content being broadcasted in subverted grupus.

For extreme conflicts, in which content removal is too weak, the protocol also
provides hard forks, as discussed in the middle of Section 2.2:

> More than simple numeric disputes, a hard fork represents a social conflict
> in which reconciling branches is no longer possible.

> The paper is too descriptive, so reading sometime is tiresome.

> Reading could be improves by using some figures to explain ideas. Sometimes
> it is difficult to follow the text.

We opted for a full description of the protocol, but we recognize that the
paper became a bit dense.
It is hard to modify the whole text without more specific suggestions.

### 2023-08-09 freechains/freechains.tex
...

@@
 
-\author{Anonymous}
-\address{Anonymous}
+\author {
+    Francisco Sant'Anna,
+    Fabio Bosisio,
+    Lucas Pires
+}
+\address {
+    UERJ - PEL - Pós-Graduação em Engenharia Eletrônica
+    \email{francisco@ime.uerj.br, fbosisio@gmail.com, lucasampires@gmail.com}
+}
 
@@
 }
+In this sense, having one or hunderds of machines in the network does not
+affect the capacity to write into the chains.
 
@@
 %Additionally,
+In our analogy with Bitcoin, Sybil attacks would require to add human resurces,
+instead of CPU power.
 
@@
 This behavior is expected to happen in the majority of the connected network.
+%
+Legitimate posts in secondary branches might be removed due to merges, which
+requires that users repost the messages.
+Nevertheless, we expect that legitimate users gossip among themselves more
+often, preventing this situation.
 %
@@
     (d) blocked messages.
+%
+The main goal of this section is to stress the consensus algorithm to show that
+permissionless public forums are viable, regardless of a slower performance in
+comparison to permissioned protocols.
 

### 2023-08-11 freechains/rebuttal.md
...

@@
 
+TODO
+
 > We do not know if the results are good or bad because there was no comparison
@@
 > The main goal of this section is to stress the consensus algorithm to show
-> that permissionless public forums are viable, regardless of a slower
-> performance in comparison to permissioned protocols.
+> that permissionless public forums are viable, regardless of the inherent
+> slower performance in comparison to permissioned protocols.
 
@@
 
-The protocol is Sybil resistant because it is expensive to write to a chain.
+In our context, a Sybil is a random machine/identity with no previous
+reputation in the chains, which, by definition, cannot operate in the chains.
 By default, users with no reps do not have write access, as described in the
@@
 
-> In this sense, having one or hunderds of machines in the network does not
-> affect the capacity to write into the chains.
+> We consider Sybils to be groups of throw-away identities and machines with
+> no previous reputation in the forums.
+> ...
+> In this sense, having one or hundreds of machines in the network does not
+> affect the capacity to write into the forums.
 > ...
 > In our analogy with Bitcoin, Sybil attacks would require to add human
-> resurces, instead of CPU power.
+> resources, instead of CPU power.
 

### 2023-08-11 freechains/freechains.tex
...

@@
 %users locked in their platforms.
-On the one hand, these companies offer free storage, friendly user interfaces,
-and robust access.
-On the other hand, they concentrate excessive power, controlling and monetizing
-over private user data.
+These companies offer free storage, friendly user interfaces, and robust access,
+but concentrate excessive power, controlling and monetizing over private user data.
 %
@@
 In centralized systems, items (i) and (ii) are trivially achieved assuming
-availability and delivery order, while item (iii) assumes that users trust the
+availability and delivery order, while item (iii) requires that users trust the
 service to moderate content.
@@
 among peers until they reach all users~\cite{p2p.survey,p2p.byz}.
-However, this approach does not guarantee consensus since posts can be received
-in conflicting orders~\cite{p2p.intention}.
+However, gossiping does not guarantee consensus since posts can be received in
+conflicting orders~\cite{p2p.intention}.
 %,p2p.dvcs}.
@@
 SPAM filters and blacklists, but which only apply a posteriori, when the
-abusive messages have already been flooded in the network.
+abusive messages have already been flooded in the network by potential
+Sybils~\cite{p2p.sybil}.
 %
@@
 \emph{multi-user/single-node consensus}, in which multiple users can exchange
-messages consistenlty within a single trusted server, but not globally across
+messages consistently within a single trusted server, but not globally across
 multiple servers.
@@
 Bitcoin~\cite{p2p.bitcoin} is the first permissionless protocol to resist
-Sybil attacks~\cite{p2p.sybil}.
-Its key insight is to rely on a scarce resource---the \emph{proof-of-work}---to
+Sybil attacks, relying on a scarce resource---the \emph{proof-of-work}---to
 establish consensus.
@@
 %
-However, Bitcoin and cryptocurrencies in general are not suitable for content
-sharing because
+However, Bitcoin and cryptocurrencies in general are not suitable for social
+content sharing because
     (i)   they enforce a unique timeline to preserve value and immunity to
@@
 eradicate Sybils, but in the context of social content sharing.
-%
-Our first contribution is to recognize the actual published contents as the
-scarce resources themselves, since they require human work.
-%
+We indicate 4 main contributions:
+%
+\begin{enumerate}
+\item Recognizing published contents as scarce resources, since they require
+human work.
 Work is manifested as new posts, which if approved by others, reward authors
@@
 %
+\item Using the reputation system to determine consensus.
 Due to decentralization and network partitions, posts in a timeline form a
@@
 %
-Our second contribution is to allow users to create diversified forums of
-interest (instead of a singleton blockchain), each counting as an independent
-timeline with its own subjective consensus etiquette.
-%
-Our third contribution is to support content removal without compromising the
-integrity of the decentralized blockchain.
+\item Allowing users to create diversified forums of interest (instead of a
+singleton blockchain), each counting as an independent timeline with its own
+subjective authoring etiquette.
+%
+\item Supporting content removal without compromising the integrity of the
+decentralized blockchain.
 Users have the power to revoke posts with dislikes, and peers are forced to
 remove payloads, only forwarding associated blockchain metadata.
+\end{enumerate}
 %
@@
 %p2p.crdts,
-To show the practicability of the consensus mechanism, we simulated months of
+To show the feasibility of the consensus mechanism, we simulated months of
 activity of a chat channel and years of a newsgroup forum, both extracted from
@@
 % and make peer-to-peer public forums practical.
+We consider Sybils to be groups of throw-away identities and machines with no
+previous reputation in the forums.
 
@@
 }
-In this sense, having one or hunderds of machines in the network does not
-affect the capacity to write into the chains.
+In this sense, having one or hundreds of machines in the network does not
+affect the capacity to write into the forums.
 
@@
 %Additionally,
-In our analogy with Bitcoin, Sybil attacks would require to add human resurces,
+In our analogy with Bitcoin, Sybil attacks would require to add human resources,
 instead of CPU power.
@@
 For instance, dislike operations in one branch to past posts in the common
-prefix could possibly block posts from users in the other branch, depending on
-how they posts are ordered.
+prefix could possibly block posts from users in the other branch.
+%, depending on how the posts are ordered.
 Therefore, we need to order branches with a global consensus, such that we can
@@
 with user $c$ because, before the forking point, $a$ and $b$ have more
-reputation than $c$, $x$, and $y$ combined.
+reputation than $c$, $x$, and $y$ have combined.
 
@@
 existed.
-As an example, suppose that the last post by $a$ (in gray) is a dislike to user
-$c$.
-Then, it's possible that the last post by $c$ (in red), now suppose with
-\nreps{0}, is rejected together with all posts in sequence, including those in
-the common suffix.
+As an example, suppose that the last post by $a$ in \code{branch-2} (in gray)
+is a dislike to user $c$.
+Then, it's possible that the last post by $c$ in \code{branch-1} (in red), now
+suppose with \nreps{0}, is rejected together with all posts in sequence,
+including those in the common suffix.
 %
@@
 identities $x$ and $y$ with the intention to accumulate \reps.
-When merging, these fake users would become legitimate and accumulate the
+When merging, these fake users would become legitimate and take over the
 majority of \reps in the forum.
@@
 %
-The same countermeasure can be used against a minority of malicious users
-colluding to boost or protect their own posts.
+%The same countermeasure can be used against a minority of malicious users
+%colluding to boost or protect their own posts.
 
@@
 disconnected for a long period of time.
-In any case, it is unacceptable that a very old remote branch affects a long
+In any case, it is unacceptable that an old remote branch affects a local
 active forum.
@@
 chance to be welcomed with a like.
-Therefore, a reasonable policy is to hold blocked posts in a temporary but
-public bag and retransmit them for some visibility.
+Therefore, a reasonable policy is to hold blocked posts in a temporary bag and
+retransmit them for some visibility.
 
@@
 the full state of the Merkle-DAG representing the causal relationships between
-the messages.
+posts.
 When peers synchronize, they mutually exchange missing branches such that they
@@
 
-The most basic concern in public forums is to resist Sybil attacks.
-Fully P2P systems cannot rely on logins or {\footnotesize CAPTCHAs} due to the
-lack of a central authority.
-Viable alternatives include (i) building social trust graphs, in which users
-already in the community vouch for new users, or (ii) imposing explicit costs
-for new posts, such as proof of work.
-%
-We propose a mix between trust graphs and economic costs.
-%
+%already in the community vouch for new users, or (ii) imposing explicit costs
+%for new posts, such as proof of work.
+%The most basic concern in public forums is to resist Sybil attacks.
+%Fully P2P systems cannot rely on logins or {\footnotesize CAPTCHAs} due to the
+%lack of a central authority.
+%Viable alternatives include (i) building social trust graphs, in which users
+%already in the community vouch for new users, or (ii) imposing explicit costs
+%for new posts, such as proof of work.
+%
+%We propose a mix between trust graphs and economic costs.
+%
+In order to resist Sybils, we propose a mix between building social trust
+graphs and imposing explicit costs for new posts:
 Rule~\code{4.a} imposes that authors require at least \onerep to post,
@@
 result of the two new consolidated posts.
-Note that rule~\code{1.b} rewards at most one post of each author at a time.
-Hence, extra posts during the 24-hour period do not reward authors with extra
-\reps.
-Note also that rule~\code{4.b} limits authors to at most \nreps{30}, which
-provides incentives to spend likes and thus decentralize the network, while
-rule~\code{4.c} limits the size of posts to at most \emph{128kB} to prevent
-DDoS attacks using gigantic blocked posts.
+%Note that rule~\code{1.b} rewards at most one post of each author at a time.
+%Hence, extra posts during the 24-hour period do not reward authors with extra
+%\reps.
+%Note also that rule~\code{4.b} limits authors to at most \nreps{30}, which
+%provides incentives to spend likes and thus decentralize the network, while
+%rule~\code{4.c} limits the size of posts to at most \emph{128kB} to prevent
+%DDoS attacks using gigantic blocked posts.
 
@@
 %
-This way, access to chains is permissionless in the sense that the actual peers
-and identities behind posts are not directly assessed by the protocol, but
-instead by the judgement of other users in the system.
+%This way, access to chains is permissionless in the sense that the actual peers
+%and identities behind posts are not directly assessed by the protocol, but
+%instead by the judgement of other users in the system.
 %
@@
 The main goal of this section is to stress the consensus algorithm to show that
-permissionless public forums are viable, regardless of a slower performance in
-comparison to permissioned protocols.
+permissionless public forums are viable, regardless of the inherent slower
+performance in comparison to permissioned protocols.
 
@@
 For each message in the original forum, we
-    (i)   set the timestamp of all peers to match the original date,
+    (i)   set the timestamp of peers to match the original date,
     (ii)  create a pair of keys if the author is new,

### 2023-08-11 freechains/freechains.tex
...

@@
 In centralized systems, items (i) and (ii) are trivially achieved assuming
-availability and delivery order, while item (iii) requires that users trust the
-service to moderate content.
+availability and delivery order, while item (iii) requires users to trust the
+service policies.
 In decentralized systems, however, none of these demands are easily
@@
 Without consensus, each client must rely on local anti-abuse policies, such as
-SPAM filters and blacklists, but which only apply a posteriori, when the
-abusive messages have already been flooded in the network by potential
-Sybils~\cite{p2p.sybil}.
+SPAM filters, but which only apply a posteriori, when the abusive messages have
+already been flooded in the network by potential Sybils~\cite{p2p.sybil}.
 %
@@
 Bitcoin~\cite{p2p.bitcoin} is the first permissionless protocol to resist
-Sybil attacks, relying on a scarce resource---the \emph{proof-of-work}---to
-establish consensus.
+Sybils, relying on a scarce resource---the \emph{proof-of-work}---to establish
+consensus.
 %The protocol maintains a single dispendious timeline consisting of linked
@@
 %
-These issues threaten our original decentralization goals, and more
-importantly, they renounce subjectivity, which is inherent to content
-moderation.
-%
-For instance, a unique timeline implies that all Internet content should be
-subject to the same consensus rules, while objective rules alone cannot
-distinguish abuse amid content.
+These issues threaten our decentralization goals, and more importantly, they
+renounce subjectivity, which is inherent to content moderation.
+%
+For instance, a unique timeline implies that all Internet content is subject
+to the same consensus rules, while objective rules alone cannot distinguish
+abuse amid content.
 %
@@
 %
-\item Allowing users to create diversified forums of interest (instead of a
-singleton blockchain), each counting as an independent timeline with its own
+\item Allowing users to create diversified forums of interest (not a
+singleton blockchain), each as an independent timeline with its own
 subjective authoring etiquette.

### 2023-08-11 freechains/rebuttal.md
...

@@
 and criticism.
-We will accommodate all clarifications and answers in a final version of the
+
+We accommodate all clarifications and answers in the final version of the
 paper.
+
+The link that follows shows the *diff* to reach the final version:
+
+- TODO
+
 Each comment is addressed individually as follows.
@@
 resto dos usuários legítimos.
-Ao ser reconectado, esse fork poderia sim ter posts removidos.
+Dessa forma, ao ser reconectado, esse fork poderia sim ter posts removidos.
 A única implicação é que ele teria que postar novamente a mesma mensagem.
@@
 
-Em termos práticos, os usuários legítimos se conectam entre si ao longo do
-tempo, o que deveria ser suficiente para evitar essa situação.
+Em termos práticos, os usuários legítimos se conectam mais diretamente ao
+longo do tempo, o que deveria ser suficiente para evitar essa situação.
 Para a versão final, nós adicionamos o seguinte trecho:
@@
 > requires that users repost the messages. Nevertheless, we expect that
-> legitimate users gossip among themselves more often, preventing this
+> legitimate users gossip among themselves more often, thus preventing this
 > situation.
@@
 
-Agora vamos discutir um possível ataque usando essa estratégia.
-O usuário malicioso poderia escrever em vários branches antes de usuários
-legítimos, de forma que os merges acabariam por remover mensagens legítimas.
-Essa situação é análoga ao comentário anterior e obrigaria os usuários a
+Como um possível ataque, o usuário malicioso poderia escrever em vários
+branches antes de usuários legítimos, de forma que os merges acabariam por
+remover mensagens legítimas.
+Essa situação, que complementa o comentário anterior, obrigaria os usuários a
 repostarem as mensagens.
-Nós entendemos que os usuários legítimos estarão mais conectados,
-inviabilizando esse tipo de ataque.
+Da mesma forma, nós entendemos que os usuários legítimos estarão mais
+conectados, inviabilizando esse tipo de ataque.
 
@@
 The main goal of our experiments in Section 3.1 is to stress the consensus
-algorithm to show that the mechanism is practical, i.e., that it has a
-reasonable performance.
+algorithm to show that the mechanism is feasible, i.e., that it has a
+reasonable performance to be used in practice.
 The good result is that the protocol works in a permissionless context,
@@
 > understand forum evolution and act accordingly. 
-
+> ...
 > I am also concerned that the content removal is too weak to work in practice.
@@
 
-This is all true, and the reason why we close the paper with your remark above.
+This is all true, and is the reason why we close the paper with your remark
+above.
 It is not even required to be "taken by extremists", but only "created by and
 for extremists".
-But we also believe that a protocol that is tamper proof and provides
-persistence can actually help law enforcement.
+But we also believe that a protocol that is tamper proof, and provides
+persistence and data transparency can actually support law enforcement.
 
@@
 > The paper is too descriptive, so reading sometime is tiresome.
-
+> ...
 > Reading could be improves by using some figures to explain ideas. Sometimes
@@
 paper became a bit dense.
-It is hard to modify the whole text without more specific suggestions.
+It is hard to modify the whole text without more specific suggestions, but we
+removed some details of Freechains from Section 3.

### 2023-08-13 freechains/freechains.tex
...

@@
 Nevertheless, we expect that legitimate users gossip among themselves more
-often, preventing this situation.
+often, thus preventing this situation.
 %
@@
 
-Many other systems have been proposed for decentralized content
-sharing~\cite{p2p.survey,p2p.ecosystem}.
-Here we consider the classes of publish-subscribe, federated, and P2P
-protocols.
+%Many other systems have been proposed for decentralized content
+%sharing~\cite{p2p.survey,p2p.ecosystem}.
+%Here we consider the classes of publish-subscribe, federated, and P2P
+%protocols.
 
@@
 mechanisms for moderation.
+
+Some works for P2P file sharing propose to account the reputation of peers to
+form a web of trust based on their behavior
+history~\cite{p2p.rep.wang,p2p.rep.eigentrust}.
+However, there are three key aspects in terms of scope that distinguishes our
+reputation system: its focus on the contents, its subjective nature, and its
+dependency on consensus.
+First, it is the actual contents that are stored and evaluated in the forums,
+with the identities being a secondary attribute.
+Second, content evaluation is subjective, as well as any decision to revoke
+posts or to fork forums.
+Third, at any given point in time, peers must agree on a common forum DAG
+prefix with global and deterministic accuracy to act identically as a whole.
+In this case, consensus is fundamental, since a subtle off-by-one discrepancy
+in reputation creates an irreconcilable fork in the network.
+Achieving consensus on a permissionless network is the key contribution of this
+work.
 

### 2023-08-14 freechains/rebuttal.md
...

@@
 
+-------------------------------------------------------------------------------
+
 > While applying the branches in order, if any post operation fails, all
@@
 > situation.
+
+-------------------------------------------------------------------------------
 
@@
 
+-------------------------------------------------------------------------------
+
 > Pela descrição, aparentemente na Figura 3 deveria ter uma transição de
@@
 
-TODO
+We added a new paragraph at the end of Section 4 (Related Work), as follows:
+
+> Some works for P2P file sharing propose to account the reputation of peers to
+> form a web of trust based on their behavior
+> history~\cite{p2p.rep.wang,p2p.rep.eigentrust}.
+> However, there are three key aspects in terms of scope that distinguishes our
+> reputation system: its focus on the contents, its subjective nature, and its
+> dependency on consensus.
+> First, it is the actual contents that are stored and evaluated in the forums,
+> with the identities being a secondary attribute.
+> Second, content evaluation is subjective, as well as any decision to revoke
+> posts or to fork forums.
+> Third, at any given point in time, peers must agree on a common forum DAG
+> prefix with global and deterministic accuracy to act identically as a whole.
+> In this case, consensus is fundamental, since a subtle off-by-one discrepancy
+> in reputation creates an irreconcilable fork in the network.
+> Achieving consensus on a permissionless network is the key contribution of this
+> work.
+
+-------------------------------------------------------------------------------
 
@@
 > slower performance in comparison to permissioned protocols.
+
+-------------------------------------------------------------------------------
 
@@
 
+-------------------------------------------------------------------------------
+
 > Maybe a way to go is to create a pool of super-reputations that could answer
@@
 
+-------------------------------------------------------------------------------
+
 > The paper is too descriptive, so reading sometime is tiresome.

### 2023-08-14 freechains/freechains.tex
...

@@
 We integrated the proposed consensus algorithm into \FC~\cite{fcs.sbseg20}, a
-practical peer-to-peer (P2P) content dissemination protocol that provides
-strong eventual consistency~\cite{p2p.sec}.
+peer-to-peer (P2P) content dissemination protocol that provides strong eventual
+consistency~\cite{p2p.sec}.
 %p2p.crdts,
-To show the feasibility of the consensus mechanism, we simulated months of
-activity of a chat channel and years of a newsgroup forum, both extracted from
-publicly available Internet archives.
+To show the feasibility of the consensus mechanism, we simulated a chat channel
+and a newsgroup forum, both extracted from publicly available Internet archives.
 
@@
 
-In the absence of moderation, permissionless P2P forums are impractical, mostly
-because of Sybils abusing the system.
-For instance, it should take a few seconds to generate thousands of fake
+Without moderation, permissionless P2P forums are impractical
+due to Sybils abusing the system.
+For instance, it should take seconds to generate thousands of fake
 identities and SPAM millions of messages.
@@
 For this reason, we propose a reputation system that works together with a
-consensus algorithm to resist Sybil attacks.
+consensus algorithm to resist Sybils.
 % and make peer-to-peer public forums practical.
-We consider Sybils to be groups of throw-away identities and machines with no
+We consider Sybils as groups of throw-away identities and machines with no
 previous reputation in the forums.
@@
 
-Note that even non-malicious network partitions may experience conflicts when
-merging branches.
+Note that even non-malicious partitions may experience conflicts when
+merging.
 For instance, dislike operations in one branch to past posts in the common
@@
 and thus \code{branch-1} is actually legitimate.
-In this case, users $a$ and $b$ might be the ones trying to take over the
-forum and would succeed, since they control the majority of \reps in the common
+In this case, users $a$ and $b$ would succeed to take over the
+forum, since they control the majority of \reps in the common
 prefix.
 Yet another possibility is that both branches are legitimate but became
-disconnected for a long period of time.
+disconnected for a long period.
 In any case, it is unacceptable that an old remote branch affects a local
@@
 For this reason, the consensus algorithm includes an extra constraint that
-prevents long-lasting local branches to merge, which leads to \emph{hard forks}
-in the network.
-A hard fork occurs when a local branch crosses a predetermined and irreversible
+prevents long-lasting local branches to merge, leading to \emph{hard forks}.
+A hard fork occurs when a local branch crosses a predetermined
 threshold of \emph{7 days} or \emph{100 posts} of activity.
@@
     \item \code{i} or \code{j} is first, whichever has more reputation in the
-          common prefix, as discussed in Section~\ref{sec.design.basic}.
+          common prefix (Section~\ref{sec.design.basic}).
     \item otherwise, branches use the arbitrary criteria of lexicographical
@@
 \FC.
-Table~\ref{fig.rules} details the concrete reputation rules we conceived for
-\FC, which is compatible with the general rules of Table~\ref{fig.general}, and
-which are discussed as follows through an example.
+Table~\ref{fig.rules} details the reputation rules we conceived for
+\FC, which is compatible with the general rules of Table~\ref{fig.general}.
+%, and which are discussed as follows through an example.
 Authors have to sign posts in order to be accounted by the reputation system
@@
 %
-In order to resist Sybils, we propose a mix between building social trust
-graphs and imposing explicit costs for new posts:
-Rule~\code{4.a} imposes that authors require at least \onerep to post,
+In order to resist Sybils, we propose a mix between social trust
+graphs and explicit costs for new posts:
+Rule~\code{4.a} imposes to authors at least \onerep to post,
 effectively blocking Sybil actions.
@@
 newbie's post to unblock it, but at the cost of \onerep.
-This cost prevents that malicious members unblock new users indiscriminately,
-which would be a breach for Sybils.
+This cost prevents malicious members to unblock new users indiscriminately.
+%which would be a breach for Sybils.
 For the same reason, rule~\code{2} imposes a temporary cost of \onerep for
@@
 %
-As detailed further, we measure the following evaluation parameters:
+As detailed next, we measure the following evaluation parameters:
     (a) metadata overhead,
@@
 %
-The main goal of this section is to stress the consensus algorithm to show that
+Our goal is to stress the consensus algorithm to show that
 permissionless public forums are viable, regardless of the inherent slower
@@
 
-Evaluation item (b) accounts the consensus algorithm applied locally at each
-chain.
-We measured the time to sort posts in a local DAG both for the first time
-(without any caches), and incrementally (from stable consensus caches).
+Evaluation item (b) accounts the consensus algorithm applied locally.
+% at each chain.
+We measured the time to sort posts in a local DAG both for the first time.
+%(without any caches),
+and incrementally from stable consensus caches.
 %
 For the chat chain, it takes \texttt{125s} and \texttt{50ms} for the initial
-and incremental sorts respectively, while for the newsgroup chain, it takes
+and incremental sorts,
+% respectively,
+while for the newsgroup chain, it takes
 \texttt{100s} and \texttt{70ms}.
@@
 Decentralized topic-based publish-subscribe protocols, such as
-    \emph{XMPP}~\cite{pubsub.xmpp},
-    \emph{ActivityPub}~\cite{pubsub.activitypub}, and
+    %\emph{XMPP}~\cite{pubsub.xmpp}, and
+    \emph{ActivityPub}~\cite{pubsub.activitypub} and
     \emph{gossipsub}~\cite{pubsub.gossipsub},
@@
 %
-In summary, \FC and \emph{pubsub} middlewares operate at different network
-layers, which suggests that \FC could benefit of the latter to manage the
-interconnections between peers.
+%In summary, \FC and \emph{pubsub} middlewares operate at different network
+%layers, which suggests that \FC could benefit of the latter to manage the
+%interconnections between peers.
 

### 2023-08-14 freechains/rebuttal.md
...

@@
 
-- TODO
+- \url{https://github.com/fsantanna-no/sbseg-23/compare/0145ec9d59ff2e1b4e1053505bf5614d63326166..b51a401cb67ece6857bd697f39df0ea428b0ba07?diff=split}
 
@@
 No meio da Seção 2.1, essa possibilidade é levantada, mas não com tanta ênfase:
-
--------------------------------------------------------------------------------
 
@@
 Em caso de empate, que pode acontecer mesmo sem usuários maliciosos, o hash do
-fork será critério considerado por todos igualmente.
+fork será o critério considerado por todos igualmente.
 Esse critério é descrito no final da Seção 2.2:
@@
 > form a web of trust based on their behavior
-> history~\cite{p2p.rep.wang,p2p.rep.eigentrust}.
+> history [Wang,EigenTrust].
 > However, there are three key aspects in terms of scope that distinguishes our
